### 2016-05-28 paper/mscov.tex
First complete draft of paper

\documentclass{article}

\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{textcomp}
\usepackage{longtable}
\usepackage{multirow, booktabs}
\usepackage{amsfonts,amsmath,amssymb}
\usepackage{natbib}
\usepackage{url}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{chemarrow}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{tikz}
\usepackage[margin=.75in]{geometry}

\renewcommand{\baselinestretch}{1.5}
\newcommand{\bl}[1]{{\mathbf #1}}
\newcommand{\bs}[1]{{\boldsymbol #1}}
\newcommand{\tr}{\text{tr}}
\newcommand{\etr}{\text{etr}}
\newcommand{\Exp}[1]{{\text{E}}[ \ensuremath{ #1 } ]  }

\begin{document}

\title{Shared Subspace Model}

\date{\today}

\title{Shared Subspace Covariance Model}
\author{Alexander Franks and Peter Hoff}
\date{\today}
\maketitle

\begin{abstract}

% From Hoff Although the covariance matrices corresponding to different
% populations are unlikely to be exactly equal they can still exhibit a
% high degree of similarity. For example, some pairs of variables may be
% positively correlated across most groups, whereas the correlation
% between other pairs may be consistently negative. In such cases much
% of the similarity across covariance matrices can be described by
% similarities in their principal axes, which are the axes that are
% defined by the eigenvectors of the covariance matrices. Estimating the
% degree of across-popu- lation eigenvector heterogeneity can be helpful
% for a variety of estimation tasks.

  We develop a model-based method for evaluating similarities and
  differences among several $p\times p$ covariance matrices in the
  $p > n$ setting.  This is done by assuming a spiked covariance model
  for each group, and sharing information about the subspace spanned
  by the group-level eigenvectors.  We propose an EM algorithm for
  estimating the shared subspace and the covariance matrix of the data
  projected onto this subspace.  We also provide a Gibbs sampling
  algorithm for estimating the posterior uncertainty of group-level
  eigenvectors and eigenvalues.  Finally, we demonstrate the utility
  of our method in an analysis of biological data.

\end{abstract}

\section{Introduction}

It has long been known that the sample covariance matrix is a poor
estimator of the population covariance matrix, particularly when the
sample size is comparable to or smaller than the number of features
($n \leq p$) \citep{Dempster1969, Stein1975}.  A host of estimators
have been devised to address this fact. These primarily include of
shrinkage based approaches \citep{Schafer2005, Ledoit2011} and low
rank factor models \citep{Mardia1980, Ullman2003, Fan2008}.  In many
cases, the covariance matrices are highly structured and thus can be
sparsely parameterized even if the matrix itself is not low rank
\citep{Burg1982, Williams1993} .  Along these lines, there has been a
recent interest in sparse Gaussian graphical model, in which the there
are assumed to be many zeros in the inverse covariance matrix
\citep{Meinshausen2006, Friedman2008}.

Most of these approaches assume that the goal is to estimate a
covariance matrix from a single dataset.  In many applications,
however, the data can be split into groups, each of which represent
samples from populations with distinct (but possibly related)
covariance matrices.  In this case, a comparative analysis of
covariance patterns across multiple groups of observations is often
warranted.  Evaluating similarities and differences between covariance
matrices can be an important complement to classical analyses which
center primarily on estimating differences in means across groups.
For instance, in case control-studies, mean level effects may be small
relative to subject variability; distributional differences between
groups may still be evident as differences in the covariances between
features.  Even when mean level differences are detectable, better
estimates of the covariability of features across groups may lead to
an improved understanding of the mechanisms underlying these apparent
mean level differences.  Further, accurate covariance estimation is an
essential part of many prediction tasks (e.g. quadratic discriminant
analysis).

By exploiting similarity across related groups of high-dimensional
data, we can increase the efficiency of existing covariance
estimators.  Some of the earliest approaches for modeling multiple
covariance matrices focused on estimation in terms of the spectral
decompositions for these matrices.  \cite{Flury1987} developed
estimation and testing procedures for the ``common principal
components'' model, in which a set of covariance matrices were assumed
to share the same eigenvectors.  \citet{Schott1991, Schott1999}
considered cases in which only certain eigenvectors are shared across
populations, and \citet{Boik2002} described an even more general model
in which eigenvectors can be shared between all or some of the groups.
More recently, \citet{Hoff2009}, noting that eigenvectors are unlikely
to be shared exactly between groups, introduced a hierarchical model
for eigenvector shrinkage based on the matrix Bingham distribution.
There has also been a significant interest in estimating covariance
matrices using Gaussian graphical models. For Gaussian graphical
models, zeros in the precision matrix correspond to conditional
independences between pairs of features given the remaining features
\citep{Meinshausen2006}.  \citet{Witten2014} extended existing work in
this area to the multi-group setting, by pooling information about the
pattern of zeros in across precision matrices.

% When the
% data can be grouped according to a set of continuous covariates
% continuously parameterized covariance models can be appropriate
% \citep{Chiu1996, Yin2010, Hoff2011}.

Another popular method for modeling relationships between
high-dimensional multivariate data related that is closely related to
low rank covariance estimation, is known partial least squares
regression (PLS). This approach, which is a special case of a bilinear
factor model, involves projecting the data into a lower dimension
space which maximizes the similarity of the two groups; this technique
does not require the data from each group to share the same feature
set.  An oft used variant for prediction, partial least squares discriminant analysis
(PLS-DA) is especially common in the chemometrics and bioinformatics
 \citep{Barker2003} .  Although closely related to
the approaches we will consider here, the primarily goal of of PLS based models
is to create regression or discrimination models, not to explicitly
infer covariance matrices from multiple groups of data.  Nevertheless,
the basic idea that data can often be well represented on a low
dimensional space is an appealing one that we leverage as well.

In many natural systems, high dimensional data is often very
structured and thus can be well understood on a relatively low
dimensional subspace. For example, with gene expression data, the
effective dimensionality is thought to scale with the number of gene
regulatory modules, not the number of genes themselves
\citep{Heimberg2016}.  As such, differences in gene expression across
groups should be expressed in terms of differences between these
regulatory modules rather than strict differences between expression
levels.  Such differences can be examined on a subspace that reflects
the correlations resulting from these modules.

Motivated by this notion, in this paper we propose a multi-group
covariance estimation model by sharing information about the subspace
spanned by group-level eigenvectors.  The shared subspace assumption
can be used to improve estimates and facilitate the interpretation of
differences between covariance matrices across groups.  For each
group, we assume ``the spiked covariance model'', a well studied
variant of the factor model \citep{Johnstone2001}.  In Section
\ref{sec:shared} we briefly review the behavior of spiked covariance
models for estimating a single covariance matrix before introducing
our extension to the multi-group setting.  In Section
\ref{sec:inference} we describe an optimization algorithm for
inferring the shared subspace and then outline a MCMC algorithm for
estimating the posterior distribution of the projected covariance
matrices.  Finally, in Section \ref{sec:simulation} we investigate the
behavior of these models in simulation before demonstrating the
utility of this model with real biological data in Section
\ref{sec:app}.

\section{The Shared Subspace Spiked Covariance Model}
\label{sec:shared}

\paragraph{The Spiked Covariance Model:}

Suppose $\bl S$ has a possibly degenerate Wishart$(\Sigma,n$)
distribution with density given by
\begin{equation} 
p(\bl S | \Sigma, n) \propto_{\Sigma} l(\Sigma:\bl S) =  |\Sigma|^{-n/2} \text{etr}( - \Sigma^{-1} \bl S/2 ) ,  
\label{eqn:lik}
\end{equation}
%
\noindent where $\Sigma \in \mathcal S_p^+$ and $n$ may be less than $p$.  Such
a likelihood results from $\bl S$ being, for example, a residual sum
of squares matrix from a multivariate regression analysis. In this
case, $n$ is the number of independent observations minus the rank of
the design matrix.  The spiked principle components model (spiked PCA)
introduced by \citet{Johnstone2001} assumes that
\begin{equation} 
\Sigma = \sigma^2 (  \bl U  \Lambda  \bl U^T  + \bl I )
\label{eqn:spiked}
\end{equation}
%
\noindent where $\Lambda$ is an $r\times r$ diagonal matrix and
$\bl U \in \mathcal V_{r,p}$, with $r \ll p$. Here $\mathcal V_{r,p}$,
known as the Stiefel manifold, comprises the set of $p \times r$
orthonormal matrices in $\mathcal{R}^p$.  The spiked covariance model
is thus a low rank factor model with homoscedastic error.  Such a
formulation is appealing because it explicitly partitions the
covariance matrix into a tractable low rank ``signal'' and isotropic
``noise''.

%  For such a covariance matrix, we have
% $\Sigma^{-1} = (\bl U \Lambda \bl U^T + \bl I )^{-1} = \bl I - \bl U
% \Omega \bl U^T $
% where $\Omega = \text{diag}(\omega_1,\ldots, \omega_r)$ with
% $\omega_s = \lambda_j/(\lambda_s+1)$.  The likelihood (\ref{eqn:lik})
% in terms of $(\sigma^2,\bl U, \Omega)$ is then
% \begin{equation}
% L(\sigma^2,\bl U , \Omega : \bl Y)  = 
% (\sigma^2)^{-np/2} \etr(-\bl S/[2\sigma^2]) \times 
%    \etr( \Omega \bl U^T [\bl S/(2\sigma^2)] \bl U ) \times  
%  \left\{ \prod_{s=1}^r (1-\omega_s)^{n/2} \right \} .  
% \label{eqn:rplik}
% \end{equation}

Classical results for parametric models (e.g., \citet{Schwartz1965})
imply that asymptotically in $n$ for fixed $p$, a Bayes estimator will
be consistent for a spiked population covariance as long as the
assumed number of spikes (non-unit eigenvalues) is greater than or
equal to the true number.  However, when $p$ is large relative to $n$,
as is the case for the examples considered here, things are more
difficult.  Under the spiked covariance model, it has been shown that
$p/n \rightarrow \gamma >0$ as $n\rightarrow \infty$, the top $r$
eigenvalues of $\bl S/(n\sigma^2)$ will converge to upwardly biased
versions of $(\lambda_1+1,\ldots, \lambda_r+1)$, if each $\lambda_s $
is greater than $\sqrt{\gamma}$ \citep{Baik2006, Paul2007} .  This has
led several authors to suggest estimating $\Sigma$ via shrinkage of
the eigenvalues of the sample covariance matrix. In particular, in the
setting that $\sigma^2$ is known, \citet{Donoho2013} propose setting
all sample eigenvalues that fall below the value
$\sigma^2(1+\sqrt{\gamma})^2$ to be $\sigma^2$, and shrinking the
larger eigenvalues in a way that depends on the particular loss
function being used.  These shrinkage functions are shown to be
asymptotically optimal in the $p/n\rightarrow \gamma$ setting.

% As demonstrated by Gavish et al \citep{Gavish2014}, even when the
% true rank is large, if blah it is still asymptotically preferable to fit a lower rank
% blah .  it is always better than hard thresholding at any other value, and is always better than ideal truncated singular value decomposition (TSVD)

Many eigenvalue shrinkage estimators can be viewed in the Bayesian
context as derivative of a Bayes estimator under a particular prior.
Note that in the absence of real prior information, invariance
considerations suggest the use of priors on $\sigma^2$ and $\bl U$
that lead to equivariant estimators.  For instance, an (improper)
prior having density $p(\sigma^2,\bl U) = 1/\sigma^2 $, with respect
to the product of Lebesgue measure on $\mathbb R^+$ and the uniform
probability measure on $\mathcal V_{r,p}$, leads to Bayes estimators
$\hat \Sigma(\bl S)$ that are equivariant with respect to rotations
and scale changes.  For multi-group covariance estimation, however, we
often have reason to believe that a priori the eigenvectors are
related across groups.  In this case, group-level equivariance to
rotations is an unreasonable assumption.  Here, both eigenvalue
\emph{and} eigenvector shrinkage can play an important role in
improving covariance estimates.  Below, we introduce our extension of
the spiked covariance model to the group setting by including an
assumption about the space shared by the span of the group-level eigenvectors.

\paragraph{The Shared Subspace Model:}

\begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.25\textwidth}
        \includegraphics[width=\textwidth]{Figs/3dplot}
        \caption{Projection in $\mathcal{R}^3$}
        \label{fig:dmelanRatio}
    \end{subfigure}
\quad
 %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.25\textwidth}
        \includegraphics[width=\textwidth]{Figs/2d-scatter}
        \caption{$Y_kV$}
        \label{fig:dmelanPosterior}
    \end{subfigure}
\quad 
    \begin{subfigure}[b]{0.25\textwidth}
        \includegraphics[width=\textwidth]{Figs/2d-scatter-orth}
        \caption{$Y_kV_{\perp}$}
        \label{fig:dmelanPosterior}
    \end{subfigure}
    \caption{ Two groups of four-dimensional data (red and blue)
      projected into different subspaces.  a) To visualize $Y_k$ we
      can project the data into $R^3$.  In this illustration, the
      distributional differences between the groups are confined to a
      two-dimensional shared subspace ($VV^T$, grey plane).  b) The
      data projected onto the two-dimensional shared subspace,
      $Y_kV$, have covariances $\psi_k$ that differ between
      groups. c) The orthogonal projection, $Y_kV_{\perp}$
      has isotropic covariance, $\sigma_k^2I$, for all groups.  }
\label{fig:shared}
\end{figure}

The spiked covariance model assumes that the goal is to estimate a
single covariance matrix.  The situation is different when we are
interested in estimating multiple covariance matrices from distinct,
but related groups.  For multi-group estimation, we can often improve
upon the efficiency of covariance estimators by pooling information
across the groups of data.  Here, we assume we have $k=1, ..., K$
groups of mean-zero normal data, $Y_k$, where $Y_k$ consists of $n_k$
independent rows of measurements on $p$ features, typically with
$n_k \ll p$.  Then, $Y_k^TY_k = \bl S_k$ follow a (degenerate) Wishart
distribution as in Equation \ref{eqn:lik}.  To improve estimation, we
seek estimators for the population covariance from one group that depend on the
data from all $K$ groups, e.g. of the form
$\hat \Sigma_k(\bl S_1, ... \bl S_K)$.  To this end, we extend the
spiked covariance model to grouped data, by assuming that the
anisotropic variability from each group can be expressed on a common
low dimensional subspace.  Specifically, we posit that the covariance
matrix for each group can be written as

\begin{equation}
\Sigma_k = V\psi_kV^T + \sigma^2_kI
\label{eqn:sspsi}
\end{equation}
%
\noindent Here, the columns of $V \in \mathcal{V}_{p, s}$ determine a subspace
shared by all groups $k$ with $p \geq s$.  Throughout this paper we
will denote to the shared subspace as $VV^T \in \mathcal G_{r,p}$, where
$\mathcal G_{r,p}$, known as the Grassmann manifold, consists of all
$s$-dimensional linear subspaces of $\mathcal{R}^p$
\citep{Chikuse2012}.  Although $V$ is only identifiable up to right
rotations, $VV^T$, which defines the plane of variation shared by all
groups, is identifiable.  Later, to emphasize the connection to the
spiked PCA model (\ref{eqn:spiked}), we will write $\psi_k$ in terms
of its eigendecomposition, $\psi_k = O_k\Lambda_kO_k$, where $O_k$ are
eigenvectors and $\Lambda_k$ are the eigenvalues of $\psi_k$ (see
Section \ref{sec:bayes}).

Note that for the shared subspace model,
$V^T\Sigma_kV = \psi_k + \sigma_k^2I$ is an anisotropic
$s$-dimensional covariance matrix for the projected data, $Y_kV$.
In contrast, the data projected onto the orthogonal space,
$Y_kV_{\perp}$, is isotropic, and typically small, for all
groups.  In Figure \ref{fig:shared} we provide a simple illustration
of this fact using simulated $4$-dimensional data from two groups.  The
differences in distribution between the groups of data can be
expressed on a two dimensional subspace spanned by the columns of
$V \in \mathcal{V}_{4, 2}$.  Differences in the correlations between
the two groups manifest themselves on this shared subspace, whereas
only the magnitude of the isotropic variability can differ between groups
on the orthogonal space.  Thus, the the shared subspace model can be
viewed as a covariance partition model, where one partition includes
the anisotropic variability from all groups and the other partition is
constrained to the isotropic variability from each group.  This
isotropic variability is often taken to be ``measurement noise''.

%Such a modelis reasonable for many applications involving groups of
%highly structure data on the same set of $p$ features.  % For instance,
% in biological applications, only certain subsets of molecules can
% physically interact which constrains the space of possible correlation
% patterns.  Our model exploits this notion, by estimating differences
% between groups on a common lower dimensional subpace.

\section{Inference}
\label{sec:inference}
In this section we outline Bayesian inference algorithms including
both an optimization algorithm and an MCMC algorithm for estimating
covariance parameters and their associated uncertainty.  MCMC
inference is particularly appealing because of the ease with which we
can quantify the uncertainty associated with each estimate.
Uncertainty quantification is particularly important in the $n \ll p$
setting where seemingly large differences in the point estimates of
the low rank covariance matrices across groups may not actually
reflect significant differences.  Although uncertainty quantification
for all parameters is desirable, for the size of the problems we are
generally interested in, we found MCMC inference for the subspace
$VV^T$ to be intractable.  Our optimization based approached is
significantly faster and works well for high-dimensional datasets.  As
such, we start, in Section \ref{sec:em}, by outlining an
expectation-maximization algorithm for estimating the shared
subspace. In Section \ref{sec:bayes} we describe a Gibbs sampler for
posterior inference of $\psi_k$ given the common subspace.  These two
algorithms can be used in conjunction to first identify and evaluate a
shared subspace and then to quantify the uncertainty associated with
the projected covariance matrices.

\subsection{Estimating the Shared Subspace}
\label{sec:em}

In this section we describe a computationally efficient algorithm for
estimating the shared subspace, $VV^T$, the projected covariance matrix
$\psi_k$ and noise parameter $\sigma_k^2$.  First, note that the full
likelihood for the shared subspace model can be written as

\begin{align}
p(S_1, ... S_k | \Sigma_k,n_k) &\propto \prod_{k=1}^K |\Sigma_k|^{-n_k/2}\etr(-\Sigma_k^{-1}\mathbf{S}_k/2)  \\
&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}\etr(-(\sigma_k^2(V\psi_kV^T +
  I))^{-1}\mathbf{S}_k/2) \\
&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}\etr(-\left[V(\psi_k +
  I)^{-1}/\sigma_k^2 V^T + (I-VV^T)/\sigma^2_k\right]\mathbf{S}_k/2)
  \\
&\propto \prod_{k=1}^K  (\sigma_k^2)^{-n_k(p-s)/2}|M_k|^{-n_k/2}\etr(-\left[VM_k^{-1}V^T + \frac{1}{\sigma^2_k} (I-VV^T)\right]\mathbf{S}_k/2) 
\end{align}
%
\noindent where we define $M_k^{-1} = (\psi_k + I) ^{-1}/\sigma_k^2$.  The log-likelihood in
$V$ (up to additive constant) is

\begin{align}
l(V) &= \sum_k tr\left(-VM_k^{-1}V^T +
       VV^T/\sigma^2_k\right)\mathbf{S}_k/2)\\
&= \sum_k \frac{1}{2}tr\left(-M_k^{-1}V^T \mathbf{S}_kV\right) + \frac{1}{2\sigma_k^2}tr\left(VV^T \mathbf{S}_k\right)
\end{align}

% \begin{align}
% p(V|...) &\propto \etr(\sum_{k=1}^K[O_k\Omega_kO^T_k]V^TS_k/(2\sigma^2_k)]V)\\
% \end{align}

We propose an expectation-maximization (EM) algorithm, to estimate the
relevant quantities.  In particular we maximize the marginal posterior
distribution of $V$, treating $M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as
the ``missing'' parameters to be imputed.  We assume independent
Jeffreys' prior distributions for both $\frac{1}{\sigma_k^2}$ and
$M_k^{-1}$ and a uniform prior for $V \in \mathcal{V}_{s, p}$.
Jeffreys' prior for these quantities corresponds to
$p(\frac{1}{\sigma_k^2}) \propto \sigma_k^2$ and
$p(M_k^{-1}) \propto |M_k^{-1}|^{-(s+1)/2}$.  From the likelihood it
can easily be shown that the conditional posterior for $M_k^{-1}$ is

$$p(M_k^{-1} | V) \propto |M_k^{-1}|^{(n_k - s -1)/2}\etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) $$

\noindent which is a Wishart($V^T\mathbf{S}_kV$, $n_k$).  The
conditional posterior distribution of $\frac{1}{\sigma_k^2}$ is simply

$$p\left(\frac{1}{\sigma^2} | V\right) \propto \left(\frac{1}{\sigma_k^2}\right)^{n_k(p-s)/2-1}\etr\left(-\frac{1}{2\sigma^2_k} (I-VV^T)\mathbf{S}_k\right)  $$

\noindent which is a Gamma($n_k(p-r)/2$,
$\frac{1}{2}tr((I-VV^T)\mathbf{S}_k)$).  Thus, for each iteration,
$t$, of the EM algorithm:

\begin{enumerate}
\item For each $k$, compute relevant conditional expectations:
\begin{enumerate}
\item $E[M_k^{-1} | V_{(t-1)}] = n_k(V_{(t-1)}^T \mathbf{S}_kV_{(t-1)})^{-1}$
\item $E[\frac{1}{2\sigma_k^2}|V_{(t-1)}] = \frac{n_k(p-r)}{tr((I-V_{(t-1)}V_{(t-1)}^T)S_k)}$
\end{enumerate}
\item Compute $V_{t}$ = $\underset{V}{argmax}  \sum_k tr\left(-VE[M^{-1}|V_{(t-1)}]^T +
       E[\frac{1}{2\sigma_k^2}|V_{(t-1)}]VV^T\right)\mathbf{S}_k/2)$ 
\end{enumerate}

For the ``M-step'', we use a numerical optimization algorithm based on
the Cayley transform to preserve the orthogonality constraints in $V$
\citep{Wen2013}.  Importantly, the complexity of this algorithm is
dominated by dimension of the shared subspace, $s$, not the number of
features $p$.  Thus, our approach is computationally efficient for
relatively small values of $s$, even when $p$ is large.

\paragraph{Evaluating a Shared Subspace:}

The above EM algorithm provides an estimate of a shared subspace, but
in practice there are no guarantees that the inferred subspace is
appropriate for all groups.  Moreover, the shared subspace can be
chosen in other ways, including more ad hoc methods involving the
singular value decomposition or simply by incorporating subject matter
considerations.  Regardless of how a particular subspace is specified,
it is important to quantify how useful this subspace is for explaining
the variability in each group.  In particular, if $V$ is a basis for
the shared subspace, then for each $k$, most of the variation in the
data should be preserved when projecting the data to this subspace.
To quantify the extent to which this is true for different groups, we
propose a simple estimator for the proportion of ``signal'' variance
that lies on any given subspace.  Specifically, for any inferred
subspace $\hat{V}$, we apply the following estimator for ratio of
  the sum of the first $s$ eigenvalues of
  $\hat{V}^T\bl \Sigma_k \hat{V}$ to the sum of the first $s$
  eigenvalues of $\Sigma_k$:
%
\begin{equation}
 \frac{||Y_k\hat{V}||_F}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{sup}
  ||Y_k\widetilde{V}||_F - \hat{\sigma}_k^2ps}
\label{eqn:ratio}
\end{equation}
%
\noindent where $||.||_F$ is the Frobenius norm. If the inferred
subspace is in fact the true subspace, i.e. $\hat{V}\hat{V}^T = VV^T$,
then this estimator will asymptotically approach one.  To see this,
note that
$\underset{V \in \mathcal{V}_{p, s}}{sup}\frac{||Y_kV||_F}{n_k}$ is
equivalent to the sum of the first $s$ eigenvalues of the sample
covariance matrix $\frac{\bl S_k}{n_k}$.  With $\hat{\lambda}^{(k)}_i$
the $i$-th eigenvalue of $\frac{\bl S_k}{n_k}$, \citet{Baik2006}
demonstrated that asymptotically as $p, n_k \rightarrow \infty$ with
$\frac{p}{n_k} = \gamma_k$ fixed

\begin{eqnarray}
\hat{\lambda}^{(k)}_i &\rightarrow& \lambda^{(k)}_i\left(1 +
                                    \frac{\sigma_k^2\gamma}{\lambda^{(k)}_i
                                    - 1}\right)\\
& \approx& \lambda^{(k)} + \sigma^2_k\frac{p}{n_k}
\end{eqnarray}

Thus,
$\sum_i^s \hat{\lambda}^{(k)}_i \approx \sum_i^s \lambda^{(k)}_i +
\sigma_k^2ps$.
If we assume that $s$ is fixed in this asymptotic regime, then
$||Y_kV||_F/n_k$ is a consistent estimator for
$\sum_i^s \lambda^{(k)}_i$.  As a consequence, the proposed estimator will be close
to one when $\hat{V}\hat{V}^T = VV^T$ and smaller if not.  The metric provides a
useful indicator for which groups can be reasonably compared on a
given subspace and which groups cannot.  In Section \ref{sec:simulation} we plot
these estimates for different simulated datasets.  In Section \ref{sec:app}
we plot the estimates for subspaces  inferred with real biological data.

\subsection{Uncertainty Quantification}
\label{sec:bayes}

Although the EM algorithm presented in the previous section yields
point estimates for $VV^T$ and $\psi_k$, it does not lead to natural
uncertainty bounds for these estimates. Full uncertainty
quantification for $VV^T$ may be desirable but we found this
challenging for the dimensions of the problems we would like to
consider.  As such, in this section, we assume that the subspace
$VV^T$ is fixed and known and focus on uncertainty quantification for
$\psi_k$ only.  Further, by focusing on group differences on a fixed
subspace, it is simpler to interpret differences across covariance
matrices.  %The shared subspace can be an estimated using the EM
% algorithm presented in Section \ref{sec:em} or chosen based on
% subject matter considerations.  
Note that when the subspace is known, the posterior distribution of
$\Sigma_k$ is conditionally independent from the other groups, i.e.
$P(\Sigma_{k} | V, \bl S_k, \Sigma_{-k}) = P(\Sigma_{k} | V, \bl
S_k)$,
so that we can independently estimate the conditional posterior
distribution for each group.

There are many different prior specifications for $\psi_k$ and
$\sigma_k^2$ that we could consider for estimating posterior
uncertainty.  One simple approach would be to use the same prior
specification used in Section \ref{sec:em}.  That is, we can use
Jeffreys' prior for $\frac{1}{\sigma_k^2}$ and
$M_k^{-1} = (\psi_k + I) ^{-1}/\sigma_k^2$.  In fact, it is easy to
show that the Gamma distribution for $\frac{1}{\sigma_k^2}$ and the
Wishart distribution for $M_k^{-1}$ are conjugate priors for the
respective quantities.  From posterior samples of $M_k$ and
$\sigma_k^2$ we can easily recover samples of $\psi_k$.
Unfortunately, if $M_k^{-1}$ has a Wishart distribution, the implied
support of $\psi_k$ is not constrained to the cone of positive
semi-definite matrices.  Direct prior specifications for $\psi_k$ are
more appealing, but many obvious choices, like the inverse-Wishart
prior distribution for $\psi_k$ do not lead to simple posterior
distributions.  As such, building on recent interest in spiked
covariance models \citep{Donoho2013, Paul2007} we focus on
tractable models based on the eigendecomposition of $\psi_k$.  Such
models are especially appealing because it is simple to summarize and
compare $\psi_k$ across groups in terms their eigenstructure (See
Section \ref{sec:simulation} and \ref{sec:app}).

By modeling the eigenstructure, we can now view each covariance
$\Sigma_k$ in terms of the spiked principle components model.
Equation \ref{eqn:sspsi}, written as a function $V$, becomes

\begin{equation}
\Sigma_k = VO_k\Lambda_KO_k^TV^T + \sigma^2_kI
\label{eqn:ss}
\end{equation}
%
\noindent Here, we make one further generalization to allow
$\psi_k = O_k\Lambda_KO_k^T$ be a rank $r \leq s$ dimensional covariance
matrix on the $s$-dimensional subspace.  Thus, $\Lambda_k$ is an $r \times
r$ diagonal matrix of eigenvalues, and $O_k \in
\mathcal{V}_{s,r}$ is the matrix of eigenvectors of
$\psi_k$.  For any individual group, this corresponds to the original spiked
PCA model (Equation \ref{eqn:spiked}) with $U_k = VO_k \in
\mathcal{V}_{p, r}$.

The likelihood for $\Sigma_k$ given the sufficient statistic
$\mathbf{S}_k = Y_kY_k^T$ is given in Equation \ref{eqn:lik}.  For the
spiked PCA formulation, we must rewrite this likelihood in terms of $V$, $O_k$,
$\Lambda_k$ and $\sigma_k^2$.  First note that by the Woodbury matrix
identity

\begin{align}
\Sigma^{-1}_k &=  (\sigma_k^2(U_k\Lambda_kU_k^T+I))^{-1}\\
&= \frac{1}{\sigma_k^2}(U_k\Lambda_kU_k^T+I)^{-1}\\
&= \frac{1}{\sigma_k^2}(I-U_k\Omega_kU_k^T)\\
\end{align}
%
\noindent where the diagonal matrix $\Omega = \Lambda(I+\Lambda)^{-1}$, e.g. $\omega_i = \frac{\lambda_i}{\lambda_{i}+1}$.  Further,

\begin{align}
|\Sigma_k| &= (\sigma_k^2)^{p}|U_k\Lambda_kU_k^T+I|\\
&= (\sigma_k^2)^{p}|\Lambda_k+I| \\
&= (\sigma_k^2)^{p}\prod_{i=1}^r(\lambda_i+1)\\
&= (\sigma_k^2)^{p}\prod_{i=1}^r(1-\omega_i)
\end{align}
%
\noindent where the second line is due to Sylvester's determinant
theorem.  Now, the likelihood of $V$, $O_k$, $\Lambda_k$ and
$\sigma_k^2$ is available from Equation \ref{eqn:lik} by substituting
the appropriate quantities for $\Sigma^{-1}_k$ and $|\Sigma_k|$ and
replacing $U_k$ with $VO_k$:

\begin{equation}
 L(\sigma_k^2,\bl V , O_k \Omega_k : \bl Y_k) \propto
    (\sigma_k^2)^{-n_kp/2}\etr(-\frac{1}{2\sigma_k^2}\mathbf{S}_k)\left(\prod_{i=1}^r(1-\omega_{ki})
   \right) ^{n_k/2}
   \etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k)
\label{eqn:sslik}
\end{equation}
%
\noindent We use conjugate and semi-conjugate priors for the parameters $O_k$,
$\sigma^2_k$ and $\Omega_k$ to facilitate inference via a Gibbs
sampling algorithm.  In the absence of real prior information,
invariance considerations suggest the use of priors that lead to
equivariant estimators.  Below we describe our choices for the prior
distributions of each parameter and the resultant conditional posterior
distributions.

\paragraph{Conditional distribution of $\sigma_k^2$:}

From Equation \ref{eqn:sslik} it is clear that the the inverse-gamma
class of prior distributions is conjugate for$\sigma_k^2$.  We chose a
default prior distribution for $\sigma^2$ that is equivariant with
respect to scale changes.  Specifically, we use Jeffreys' prior, an
improper prior with density $p(\sigma^2) \propto 1/\sigma^2 $.  Under
this prior, straightforward calculations show that the full
conditional distribution of $\sigma_k^2$ is
inverse-gamma$( n_k p/2 , \tr(\bl S_k(\bl I -\bl U_k\Omega_k\bl
U_k)/2)$, where $U_k = VO_k$

\paragraph{Conditional distribution of $O_k$:} Given the likelihood
from Equation \ref{eqn:sslik}, it is easy to show that the class of
von Mises-Fisher-Bingham distributions are conjugate for $O_k$
\citep{Hoff2009, Hoff2012}.  Again, invariance considerations
lead us to use a rotationally invariant uniform probability measure on
$\mathcal V_{s,p}$.  Under this uniform prior, the full conditional
distribution of $\bl O_k$ has a density proportional to the
likelihood, so

\begin{align}
\label{lik_vo}
 p(O_k | \sigma^2_k, U_k, \Omega_k) & \propto \etr(\Omega_kO^T_kV^T[S_k/(2\sigma^2_k)]VO_k)
\end{align}
%
\noindent This is a Bingham$(\Omega, \hat{V}^T \bl S_k \hat{V}/(2\sigma^2)$
distribution on $\mathcal V_{s, r}$ \citep{Khatri1977}. A
Gibbs sampler to simulate from this distribution is given in
\citet{Hoff2012}.

Together, the prior for $\sigma_k^2$ and $O_k$ leads to conditional
(on $V$) Bayes estimators $\hat \Sigma(V^T \bl S V)$ that are
equivariant with respect to scale changes and rotations on the
subspace spanned by $V$, so
that $\hat \Sigma(a \bl W V^T \bl S V \bl W^T) = a \bl W \hat\Sigma(V^T
\bl S V) \bl W$
for all $a>0$ and $\bl W\in \mathcal O_{s}$ (assuming an invariant
loss function). Interestingly, if $\Omega$ were known (which it is
not), then for a given invariant loss function the Bayes estimator
under this prior minimizes the (frequentist) risk among all
equivariant estimators \citep{Eaton1989}.

\paragraph{Conditional distribution for $\Omega$:} Here we specify the conditional
distribution of the diagonal matrix $\Omega_k =
\Lambda_k(I+\Lambda_k)^{-1} = \text{diag}(\omega_{k1}, ... \omega_{kr})$.  We first consider a uniform(0,1) prior distribution for each element of $\Omega$, or
equivalently, an $F_{2,2}$ prior distribution for the elements of
$\Lambda$.  The full conditional distribution of an
element $\omega_i$ of $\Omega$ is proportional to the likelihood
function, so

\begin{align}
p(\omega_{ki}|V, O_k, \bl S_k) &\propto_{\omega_{ki}}
  \left(\prod_{i=1}^r(1-\omega_{ki})^{n_k/2}  \right)
  \etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k) \\
&  \propto  (1-\omega_{ki})^{n/2} e^{c \omega_s  n/2},    
\label{eqn:wpost}
\end{align}
%
\noindent where $c = \bl u_{ki}^T S_k \bl u_{ki}/(n \sigma^2)$ and $\bl u_{ki}$ is
column $i$ of $U_k = VO_k$.  While not proportional to a density
belonging to a standard class of distributions, we can sample from the
corresponding univariate distribution numerically.  The behavior of
this distribution is straightforward to understand: if $c\leq 1$, then
the the function has a maximum at $\omega_i =0$, and decays
monotonically to zero as $\omega \rightarrow 1$.  If $c>1$ then the
function is uniquely maximized at $(c-1)/c \in (0,1)$.  To see why
this makes sense, note that the likelihood is maximized when the
columns of $\bl U_k$ are equal to the eigenvectors of $\bl S_k$
corresponding to its top $r$ eigenvalues
\citep{Tipping1999}. At this value of $\bl U_k$, $c$ will then
equal one of the top $r$ eigenvalues of $\bl S_k/(n\sigma_k^2)$.  In the
case that $n\gg p$, we expect
$\bl S_k/(n_k\sigma_k^2)\approx \Sigma_k/\sigma_k^2$, the true (scaled)
population covariance, and so we expect $c$ to be near one of the top
$r$ eigenvalues of $\Sigma_0/\sigma^2$, say $\lambda_0+1$.  If indeed
$\Sigma_k$ has $r$ spikes, then $\lambda_k>0$,
$c \approx \lambda_0 +1 > 1$, and so the conditional mode of $w_i$ is
approximately $(c-1)/c = \lambda_i/(\lambda_i+1)$, the correct value.
On the other hand, if we have assumed the existence of a spike when
there is none, then $\lambda_0=0$, $c\approx 1$ and the Bayes estimate
of $w_i$ will be shrunk towards zero, as it should be.

% \paragraph{Conditional distribution of $V$}
% Need to finish / should we include?

% Although we prefer EM... it is also posible to estimate $V$ using
% Bayesian inference.  We assume a uniform prior on $V \in \mathcal{V}_{p, s}$ so that the
% conditional posterior of $V$ is proportional to the likelihood:

% \begin{align}
% p(V|\bl O, \bl S, \bl \sigma^2, \bl \Omega) &\propto \etr(\sum_{k=1}^K[S_k/(2\sigma^2_k)]V[O_k\Omega_kO^T_k]V^T)\\
% \end{align}

% Although close in appearance, this conditional distribution is not a matrix
% Bingham-von Mises-Fisher distribution.  However, the conditional distribution
% of a column of $V$ given all other columns is a vector Bingham-von
% Mises-Fisher distribution.  To see this, let
% $A_k = S_k/(2\sigma^2_k)$ and $B_k = O_k\Omega_kO^T_k$.  We can write
% the conditional distribution of $V$ given $A_k$ and $B_k$ as
% \begin{align}
% P(V|A_k,B_k) &\propto \etr(\sum_{k=1}^KA_kVB_kV^T)\\
% &= \etr(\sum_{k=1}^K\sum_{i,j}^sA_kv_iv_j^Tb_{ijk})\\
% &= \etr(\sum_{i,j}^s \sum_{k=1}^K v_j^T[b_{ijk}A_k]v_i)
% \end{align}

% From here, we specify the conditional distribution of a single column of $V$ given $A_k$, $B_k$ and all other columns of V:

% \begin{align}
% P(v_i|A_k,B_kV_{-i})  &\propto \etr(2(\sum_{j\neq i}^s v_j^T[\sum_{k=1}^K b_{ijk}A_k])v_i+v_i^T[\sum_{k=1}^K b_{ijk}A_k]v_i))\\
% &= \etr(Cv_i+v_i^TMv_i)
% \label{eqn:vbmf}
% \end{align}

% This is a vector Bingham-von Mises-Fisher Distribution, BMF(C, M)
% with $C = 2(\sum_{j\neq i}^s v_j^T[\sum_{k=1}^K b_{ijk}A_k])$ and $M=
% \sum_{k=1}^K b_{ijk}A_k$.  Following \citet{Hoff2012},  we can derive
% a Gibbs sampler for $V$ based on conditional draws of the columns of
% $V$.

\section{Simulation Study}
\label{sec:simulation}

\subsection{Validation of Inference}

In this section we start by validating, in simulation, that our
inferential algorithm recovers the true parameters of the data
generating mechanism.  We simulate five groups of $p=200$ dimensional
data each with sample size $n_k=50$.  We choose the eigenvectors of
the anisotropic component of the covariance matrix, $\psi_k$, to share
a common two dimensional subspace.  To fit the model to this simulated
data, we first estimate the shared subspace using the EM algorithm outlined
in Section \ref{sec:em} and then use the MCMC algorithm described in
Section \ref{sec:bayes} to estimate the posterior distribution of the
projected covariance matrices $\psi_k$ and noise variance $\sigma_k^2$.

\begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{Figs/posteriorRegions}
        \caption{Posterior}
        \label{fig:simPosterior}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{Figs/simRatio-05-19}
        \caption{Ratio}
        \label{fig:simRatio}
    \end{subfigure}
    \caption{a) 95\% posterior regions for the ratio of the
      eigenvalues, $\frac{\lambda_1}{\lambda_2}$, of $\psi_k$ and the
      orientation of the principle axis on space spanned by $\hat{V}$
      cover the truth.  True parameter values on
      $\hat{V}^T\Sigma_k\hat{V}$ are indicated by the crosses.  Since
      $V$ is only identifiable up to rotation we find the Procrustes
      rotation that maximizes the similarity of $\hat{V}$ to the true
      data generating basis. True eigenvalue ratios were 10 (red and
      black), 3 (green and blue) and 1 (light blue).  True
      orientations were $\pi/4$ (black), $-\pi/4$ (red and green) and
      0 (blue and cyan).  b) Estimated proportion of variance
      explained by the subspace (Equation \ref{eqn:ratio}).  As
      expected, for each group, the inferred subspace explains
      essentially all of the variance from the first two eigenvectors of
      $\Sigma_k$.  }
\end{figure}

For this two dimensional shared subspace model we summarize the
estimates of $\psi_k$ in terms of its eigendecomposition.
Specifically, we compute the eigenvalue ratio
$\frac{\lambda_1}{\lambda_2}$, with $\lambda_1 > \lambda_2$, and the
angle of the first eigenvector on this subspace,
$\text{arctan}(\frac{O_{12}}{O_{21}})$, relative to the first column
of of $V$.  Figure \ref{fig:simPosterior} depicts the 95\% posterior
regions for these quantities.  We select the displayed posterior
samples in each group by iteratively remove the points corresponding
to the vertices of the convex hull until only 95\% of the original
samples remain.  Non-overlapping posterior regions are indicative of
significant differences in the covariance structure between groups.

In this example, the ratio of the eigenvalues of the true covariance
matrices are $10$ (red and black groups), $3$ (green and blue
groups) and $1$ (cyan groups).  Larger values of this ratio
correspond to more correlated contours and a value of $1$ implies
isotropic covariance.  The posterior regions for all groups cover the
true eigenvalue ratio and primary axis orientation of $\hat{V}^T\Sigma_k\hat{V}$.
Note that for smaller eigenvalue ratios, there is larger uncertainty
about the orientation of the primary axis.  When the ratio is one, as
is the case for the cyan colored group, there is no information
about orientation of the primary axis since the contours are
spherical.

We also verify that the data projected onto the inferred subspace
preserves most of the variability in the data.  To do this, we compute
estimates for the proportion of explained variance, as given in
Equation \ref{eqn:ratio}.  Figure \ref{fig:simRatio} depicts a barplot
of this estimate for each group.  For each group, the estimate is
nearly one, which reflects the fact that we were able to infer a two
dimensional shared subspace that explains most of the variance from
the first two eigenvectors of $\Sigma_k$.  In this example, by
construction we ensured that such a subspace exists. However, this is
not always the case.  In the next section we investigate examples for
which the inferred subspace may not capture the relevant variation for
all groups.

\subsection{Rank Selection and Model Misspecification}

\begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{Figs/LossVsDimension}
        \caption{Stein's Loss vs s}
        \label{fig:sdimension}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{Figs/simRatio-s5}
        \caption{s = 5}
        \label{fig:ratio-s5}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{Figs/simRatio-s20}
        \caption{s = 20}
        \label{fig:ratio-s20}
    \end{subfigure}
    \caption{a) Average Stein's loss as a function of the dimension of
      the shared subspace dimension.  Data from ten groups, $U_k$
      uniform on the Stiefel manifold $\mathcal{V}_{200, 2}$.  As the
      shared subspace dimension increases, $s \rightarrow p$, the loss
      converges to the loss from the independent spiked covariance
      estimates (dashed line). b) For a simulated dataset, estimated
      fraction of variance explained when the assumed shared subspace
      is dimension 5.  c).  For a single simulated dataset, estimated
      fraction of variance explained when the assumed shared subspace
      is dimension 20.  We can capture nearly all of the variability
      in each of the 10 groups using an $s=rK=20$ dimensional shared
      subspace. }
\label{fig:dimensionPlots}
\end{figure}

Naturally, shared subspace inference works well when the model is
correctly specified.  What happens when the model is not well
specified?  We explore this question in silico by simulating data from
different data generating models and evaluating the efficiency of
various covariance estimators.  In all of the following simulations we
evaluate covariance estimates using Stein's loss,
$L_S( \Sigma_k , \hat\Sigma_k) = \text{tr}( \Sigma_k^{-1} \hat
\Sigma_k ) - \log |\Sigma_k^{-1} \Sigma_k | - p$.
Since we compute estimates for multiple groups of data, we report the
average Stein's loss
$L(\Sigma_1, ..., \Sigma_K; \hat\Sigma_1, ..., \hat\Sigma_K ) =
\frac{1}{K} \sum_k L_S( \Sigma_k , \hat\Sigma_k)$.
Note that under Stein's loss, the Bayes estimator is the inverse of
the posterior mean of the precision matrix,
$\hat \Sigma_{k} = \Exp{ \Sigma_k^{-1} | \bl S_k}^{-1}$ which we
estimate using MCMC samples.

We start by investigating the behavior of our model when we
underestimate the true dimension of the shared subspace.  In this
simulation, we generate data from five groups of mean-zero normally
distributed data in a $p=200$ dimensional feature space.  For each
group we use the spiked covariance model of Equation \ref{eqn:spiked},
where we generate eigenvectors uniformly from the Stiefel manifold,
$U_k \in \mathcal{V}_{200, 2}$ with eigenvalues
($\lambda_1, \lambda_2) = (250, 25)$ .  In this case, although most of
the variation in each group can be captured by a two dimensional
subspace, these are not the same subspaces across groups.  We make no
assumption, a priori, that the subspaces are similar across groups.
Below, we evaluate how well the shared subspace estimator performs with such
data for inferred subspaces of increasing dimension.

In Figure \ref{fig:sdimension} we plot the average Stein's loss, for
estimates derived from a shared subspace model for different values
of the shared subspace dimension, $s$.  For each value of $s$, we
report the loss averaged over ten independent simulations.  In this
figure, the dashed blue line corresponds to the average Stein's loss
for covariance matrices estimated independently, e.g. without the
shared subspace assumption.  As the shared subspace dimension increases to the
dimension of the feature space, $s \rightarrow p$, the loss for the
shared subspace estimator quickly converges to the loss for the
individual spiked covariance estimator.  Note that
it is always true that span($[U_1, ..., U_K]) \leq rK$.  Even when there is little
similarity between the eigenvectors from each group, the shared
subspace estimator with $s = rK$ will perform comparably to the
estimator, provided that we can identify a subspace, $\hat{V}\hat{V}^T$, where
span($\hat{V})$ is close to $\text{span}([U_1, ..., U_K])$.  This is
immediately apparent in Figure \ref{fig:sdimension} from the phase transition
in Stein's loss at $s = 20$.

From the above simulation, it is clear that correctly specifying the
dimension of the shared subspace is important for efficient covariance
estimation.  When the dimension of the shared subspace is too small,
there is larger loss for estimate relative to the truth.  We can also
use the available data, independent of the true covariance matrices,
to evaluate the appropriateness of the inferred subspace.  To this
end, we calculate the estimate of the ``proportion of explained
variance'' estimator from Section \ref{sec:em} (Equation
\ref{eqn:ratio}).

Figure \ref{fig:dimensionPlots}b and c depict the estimated proportion of
variance explained for a single dataset for each of five group, using
inferred subspaces of two different dimensions.  For the results
displayed in Figure \ref{fig:ratio-s5}, we fixed the dimension of the
subspace to $s=5$ (clearly too small) and for the results displayed in
Figure \ref{fig:ratio-s5} we fixed the dimension of the shared subspace
to $s=20$.  When the shared subspace model is appropriate, the
subspace spanned by $\hat{V}$ explains most of the variance in
$\Sigma_k$, for all groups.  When the estimated subspace is too small,
the proportion of variance will be smaller than one for at least some of the
groups.  For a given fixed subspace, these plots can be used to decide
which groups can be fairly be compared on this space and which groups
cannot and whether we would benefit from fitting a model with
increased dimension $s$.

\paragraph{Model Comparison and Rank Estimation:}

In the previous section, it was apparent that correct specification
for the rank of the shared subspace is important for efficient
inference.  So far, we have assumed that the group rank, $r$, and
shared subspace dimension, $s$, are fixed and known, but in practice
this is not the case.  Prior to fitting a model we should estimate
these quantities.  \citet{Gavish2014} provide an asymptotically
optimal (in mean squared error) singular value threshold for low rank
matrix recovery with noisy data.  We use their rank estimator, which
is a function of the median singular value of the data matrix and the
ratio $\gamma_k =\frac{p}{n_k}$, to estimate the rank, $r$, of a
spiked covariance matrix.  Although their estimator was derived for
individual covariance estimation, we found as a useful heuristic,
Gavish and Donoho's estimator can also be applied to the pooled data
to estimate the shared subspace dimension, $s$.  Specifically, we concatenate
the data from all groups to create an $(\sum_k n_k) \times p$
dimensional matrix and apply their rank estimator to this matrix to
choose $s$.

Using these rank estimators, we conduct a simulation which demonstrates
the relative performance of shared subspace group covariance
estimation under different data generating models.  In this
simulation, we consider three different spiked covariance data models:
1) ``complete pooling'', in which the spiked covariance matrices from
all groups are identical, e.g.
$\Sigma_1 = \Sigma_2 = ... \Sigma_K = U\Lambda U^T + \sigma^2I$; 2)
the ``shared subspace'' model, outlined in Section \ref{sec:shared};
and 3) ``no pooling'', e.g. the data from each group is generated from
a spiked covariance matrix with each $U_k$ drawn as an independent
uniform sample from the Stiefel manifold $\mathcal{V}_{r, p}$.  For
each of these data generating models, we simulate datasets with
$p=200$ features, rank $r=2$ and eigenvalues
$(\lambda_1, \lambda_2) = (250, 25)$.  For data generated under the
shared subspace model, we additionally assume that the two
eigenvectors of $\psi_k$ span a common two dimensional subspace, e.g. we fix
$s=2$.  We simulate 100 independent datasets for each data generating
mechanism.

For each data generating mechanism we fit each of three corresponding
inferential models (``complete pooling'', ``shared subspace'' and ``no
pooling''). For the complete pooling and no pooling models, which
involve independent estimation of individual covariance matrices, we
estimate the rank $r$, and then proceed with Bayesian inference of the
eigenvectors, eigenvalues, and noise variance.  For the shared
subspace estimator, we first estimate the shared subspace dimension
$s$ and then infer the optimal $s$-dimensional shared subspace,
$\hat{V}\hat{V}^T$, using the EM algorithm presented in Section
\ref{sec:em}.  We then use Bayesian inference to estimate the
posterior distribution of the projected covariance matrices
conditional on $\hat{V}\hat{V}^T$.  In Table \ref{table:groupLoss} we report the
average Stein's loss, with corresponding 95\% intervals, for the
estimates derived from each inferential model.

\begin{table}
\begin{center}
  \begin{tabular}{ l  l | c | c | c |}
    \multicolumn{2}{c}{} & \multicolumn{3}{c}{\textbf{Inferential Model}} \\
  \multicolumn{2}{c|}{}  & Shared Subspace & Complete Pooling & No pooling \\  \cline{2-5}
    \multirow{3}{*}{\rotatebox[origin=c]{90}{\textbf{Data Model}}} & 
   Shared Subspace & 0.8 (0.7, 0.9) & 2.1 (1.7, 2.6) & 3.0 (2.9, 3.2) \\ \cline{2-5}
   & Complete Pooling & 0.8 (0.7, 0.9) & 0.7 (0.6, 0.8) & 3.0 (2.9, 3.2)\\ \cline{2-5}
   & No pooling & 7.1 (6.2, 8.0) & 138.2 (119, 153) & 3.0 (2.9, 3.2) \\ \cline{2-5}
  \end{tabular}
  \caption[Table caption text]{Stein's loss (and 95\% intervals) for
    different inferential and data generating models.  For each of
    $K=10$ groups, we simulate a $p=200$ dimensional spiked covariance
    matrix for rank $r=2$.  For data from a shared subspace model the
    eigenvectors of the covariance matrix share a subspace dimension
    $s=2$.  All results are based on $n_k = 50$ observations per
    group.  }
\label{table:groupLoss}
\end{center}
\end{table}

As expected, the estimates with the lowest loss are derived from the
inferential model that matches the data generating model. However, the
shared subspace estimator has small loss under model misspecification
relative to the alternative.  For example, when the data is generated
using the complete pooling model, the shared subspace estimator has
almost four times smaller loss than the no pooling estimator.  When
the data is generated under the no pooling model, the shared subspace
estimator is over an order of magnitude better than the complete
pooling estimator.  Thus, shared subspace estimation can be an
appropriate choice when the amount of similarity between groups is not
known a priori.

The shared subspace model combined with the rank estimation procedure
suggested by \citet{Gavish2014} can be widely applied to group
covariance estimation because the estimator adapts to the amount of
similarity between groups.  Importantly, the shared subspace model can
be viewed as a generalization of the standard ``no pooling'' spiked covariance
model. The two are equivalent when we let the shared subspace span the
full space, e.g. $V \in \mathcal{O}(p)$.  As shown in Figure
\ref{fig:sdimension}, the shared subspace model performs nearly as
well as the no pooling estimator when the subspace dimension is equal to $s = rK$.
In the simulations used to create Table \ref{table:groupLoss}, we
typically estimated $s$ to be slightly smaller than $rK$ for the ``no
pooling'' data, which explains the slightly higher loss for the shared
subspace model. With improved estimators for $s$, the shared subspace
model should be even more competitive.

% Ultimately, in many real world examples, we are more interested in
% finding a useful space to interpret and compare differences between
% covariance matrices than we are in finding the lowest-loss estimator
% for the full covariance matrices.  In this context, the relevant
% quantity is often an estimate of the projected covariance, $V^T\Sigma
% V$.

% In Figure \ref{fig:lossBoxplot} we depict boxplots for
% the distribution of losses,
% $L_S(\hat{V}^T\Sigma_k\hat{V}, \hat{V}^T\hat{\Sigma}_k\hat{V})$, on
% simulated data.  In the left boxplot the data come from a
% 2-dimensional shared subspace model and on the right plot the data
% come from a 10-dimensional shared subspace model.  For both boxplots,
% estimates $\hat{V}^T\hat{\Sigma}_k\hat{V}$ are derived from assumed
% 2-dimensional model. This figure demonstrates that the loss for
% estimating the projected covariance matrix is not significantly larger
% when we underestimate the dimension of the true shared subspace.

\section{Analysis of Biological Data}
\label{sec:app}

In this Section, we apply the shared subspace spiked covariance model to data in two
different biological examples.  In the first, we compare gene
expression data from juveniles with different subtypes of leukemia.  Even
after removing mean effects, the data indicate significant differences
in covariance matrices between groups.  In the second example, we
compare covariance matrices in a metabolomic analysis of fly data.  In
this example, after controlling for mean effects due to fly age, gender and
genotype, we find little compelling evidence of differences in metabolite
covariances between groups.

In these analyses we employ a variation of the shared
subspace model.  Specifically, we define the $s$ by $s$ matrix $\psi_k$ as

\begin{equation}
\psi_k =\left( \begin{array}{cc}
O_k\Lambda_kO_k^T & 0  \\
0 & D  \end{array} \right)
\end{equation}

For the following analyses $O_k\Lambda_kO_k^T$ is a rank $r=2$ matrix
and $D$ is an $(s-r)$-dimensional diagonal matrix.  We write
$V = [V_1\, V_2]$, with $V_1 \in \mathcal{V}_{p,r}$ as the basis for a
$r$-dimensional shared subspace that explains the differences between
groups and $V_2 \in \mathcal{V}_{p,(s-r)}$ corresponding to the remaining $(s-r)$
eigenvectors common to all groups.  Importantly, we make no assumptions about the
magnitude of the eigenvalues $\Lambda_k$ relative to the eigenvalues
$D$, so that the largest eigenvectors of $\psi_k$ may correspond to
the eigenvectors $V_2$.

We find this formulation useful because in real world analyses,
differences between groups may not manifest themselves in the first
principle components.  For instance, in genetic analyses of humans,
there may be large person to person variability common to all groups
which is unrelated to the grouping of those subjects. Differences
between groups may still manifest themselves in the smaller principle
components. This formulation allows complete pooling of most
eigenvectors, yet a space to identify and compare relevant
differences.  Moreover, visualizing and interpreting differences
between groups on higher dimensional spaces is a challenging problem.
Using this framework we can easily understand the largest differences
across groups in a two dimensional space while completely pooling
remaining anisotropic variability among the groups.

\paragraph{Analysis of Gene Expression Data}

Here, we demonstrate the utility of the shared subspace covariance
estimator for exploring differences in the covariability of gene
expression levels in young adults with different subtypes of pediatric
acute lymphoblastic leukemia \citep{Yeoh2002}.  The raw data consists
of gene expression levels for over 12,000 genes in seven different
subtypes of leukemia: BCR-ABL, E2A-PBX1, Hyperdip50, MLL, T-ALL,
TEL-AML1 and a seventh group for unidentified subtypes.  The groups
have corresponding sample sizes of $n = (15, 27, 64, 20, 43, 79, 79)$.
Although there are over 12,000 genes, the vast majority of expression
levels are missing.  Thus, we restrict our attention to the genes for
which less than half of the values are missing and use Amelia, a
software package for missing value imputation, to fill in the
remaining missing values \citep{Amelia}.  After restricting the data
in this way, $p=3124$ genes remain.  Prior to analysis, we demean both
the rows and columns of the gene expression levels in each group.

In Figure \ref{fig:leukemia} we plot the results of our analysis.
Panel \ref{fig:leukemiaRatio} shows that over $40\%$ of the variance
can be explained by the estimated two-dimensional subspace for all groups, with 4 of
the 6 groups over 60\%.  Panel \ref{fig:leukemiaPosterior} reflects some significant
differences in the posterior distribution of eigenvalues and
eigenvectors between groups.  The $x$-axis corresponds to the
orientation of the principle axis of $\psi_k$ on $V$ and the $y$-axis
corresponds to the ratio of the larger eigenvalue over the smaller
eigenvalue.  Several groups show significantly different orientations
for the first principle component of the subspace $\hat{V}$.  Further,
the subtype E2A-PBX1 has a larger eigenvalue ratio which reflects a
more correlated distribution on the subspace spanned by $V$.  The
TEL-AML1 and Hyperdip50 appear to cluster together which suggests
there is little detectable difference between their covariance
structure.  Note that when the eigenvalue ratio is close to one (as is
the case for the ``other'' subtype), the distribution of expression
values on the subspace is nearly spherical and thus the orientation of
the primary eigenvector is at best weakly identifiable.  This is
reflected in the wide posterior range of orientations of the first
principle component.

Further intuition about the differences in covariances between groups
can be understood in the biplot in Figure \ref{fig:leukemiaBiplot}.
Here, we plot the contours of the covariance matrices for three of
subtypes of Leukemia.  We also plot different shapes for the 1\% of genes with
the largest loadings on the first two columns of $V$ and the
remaining values as light grey dots.  The genes with the largest
loadings are clustered by quadrant and listed in the corresponding
table.  Even though we remove all mean-level differences in gene
expressions between groups, we are still able to identify genes with
known connections to cancer and leukemia.  As one example, MYC (top
right) is an oncogene with well established association to many
cancers \citep{Dang2012}.

\begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{Figs/leukemiaRatio-04-27}
        \caption{Proportion of Variance on $\hat{V}$}
        \label{fig:leukemiaRatio}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{Figs/leukemiaPosterior-04-27}
        \caption{95\% Posterior Regions }
        \label{fig:leukemiaPosterior}
    \end{subfigure}
\caption{}
\label{fig:leukemia}
\end{figure}

In this figure, genes that fall in the upper right quadrant have
positive loadings in both columns of $V$. Genes in the upper left
corner of the figure have positive loadings on the first column of $V$
but negative loadings on the second column of $V$.  The principle axis
for a group aligns in the direction of genes that exhibit the largest
variability in that group.  Genes which lie in a direction orthogonal
to the principle axis exhibit reduced variability, relative to the
other groups. For instance, genes in the upper left corner of this
plot (triangles) exhibit large, positively correlated variability in
the Hyperdip50 group.  In this same group, there is reduced
variability, relative to the other groups, among the genes in the
upper right corner of the plot (squares), since this cluster of genes lie in
a direction orthogonal to the principle axis.  In contrast, the
BCR-ABL group is aligned primarily with the second column of $V$,
which means that the genes indicated by squares and circles vary
significantly in this group.  Genes in the square group are
anti-correlated with those in the circle group, since the their
loadings have opposite signs on $V2$.

  \begin{figure}[!ht]
    \centering
    \includegraphics[width=0.5\textwidth]{Figs/leukemia-biplot-05-13}
    \qquad
\raisebox{1.25\height}{
\footnotesize
\input{Figs/biplot.tab}
}
\caption{Left) Variant of a biplot with contours for three leukemia
  subtypes and the loadings for each gene on the first two columns of
  $\hat{V}$.  All gene's loadings are displayed in light gray, and the
  top 1\% of genes with with the largest magnitude loadings are
  displayed as either a triangle, square or circle depending on which
  quadrant they lie in.  Right) List of the gene's with the largest
  loadings, grouped by quadrant. }
\label{fig:leukemiaBiplot}
  \end{figure}

\paragraph{Analysis of Metabolomic Data}

Next, we apply shared subspace group covariance estimation to a
metabolomic data analysis on fly aging \citep{Hoffman2014}.  We bin
the data to include groups of flies less than 10 days (young), between
10 and 51 days (middle) and greater than 51 days (old), and further
split by gender. The sample sizes range from 43 to 59 flies per group.
We analyze metabolomic data corresponding to metabolites with 3714
mass-charge ratios.  After removing mean effects due to age and gender
we fit the shared subspace estimator and compare covariances between
groups.  Interestingly, we identify a subspace that explains almost
all of the variability in the first few components of $\Sigma_k$ for
all groups (Figure \ref{fig:dmelanRatio}). However, we see little
evidence for differences in metabolite covariances on this subspace,
as evidenced by the large overlap in posterior distributions for each group
(Figure \ref{fig:dmelanPosterior}). This is indicative of large
amount of variation which is common across all groups for the first
$s$ principle components. Differences in covariation across may
manifest themselves for the smaller principle components, but are too
small to detect given the sample sizes.

\begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{Figs/dmelanRatio-04-28}
        \caption{Ratio}
        \label{fig:dmelanRatio}
      \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.45\textwidth}r
        \includegraphics[width=\textwidth]{Figs/dmelanPosterior-04-28}
        \caption{Posterior}
        \label{fig:dmelanPosterior}
    \end{subfigure}
    \caption{a) Proportion of variance on $\hat{V}$ is close to one
      for all groups suggesting that blah. However, there is
      significant overlap in the posterior distribution of $\Lambda_k$
      and $U_k$. This suggests that there is little difference in the
      covariance matrices between groups, on this particular subspace.  }
\end{figure}

\section{Discussion}

In this paper, we proposed a hierarchical model for estimating and
comparing differences in covariance matrices across multiple groups on
a common low dimensional subspace.  We described an EM algorithm for
estimating this common subspace and a MCMC algorithm for inferring the
projected covariance matrices and their associated uncertainty.
Estimates of both the shared subspace and the projected covariance
matrices can be useful summaries of the data.  For example, with the
biological data, the shared subspace highlights the full set of genes
or molecules that might be correlated in any group.  Differences between
group covariance matrices can be understood in terms of differences in
these sets of correlated and anti-correlated genes.  In this applied
analysis, we demonstrated how we can use these notions to visualize
and contrast the posterior distributions of covariance matrices
projected onto a particular subspace.

In simulation, we showed that the shared subspace model can
still be a reasonable choice for modeling multi-group covariance
matrices even when we have little knowledge, a priori, of how similar
the groups are.  Even when there is little to no similarity, the shared
subspace model is appropriate, as long as the dimension of the shared
subspace is large enough.  However, selecting the rank of the shared
subspace remains a significant practical challenge.  Although we
propose a useful heuristic for choosing the dimension of the shared
subspace based on the rank selection estimators of \citet{Gavish2014},
a more principled approach is warranted.

It was also a challenge to estimate the subspace $VV^T$ once we had
chosen a rank.  In our approach, we estimate the shared subspace
$VV^T$ separately from the MCMC inference of $\psi_k$ because we found
estimation of $VV^T$ using MCMC inference to be
particularly difficult.  Although fixing the shared subspace
simplifies interpretation of the results, in many cases it may be
desirable to develop a fully Bayesian approach for jointly estimating
$VV^T$ and $\psi_k$.  Recent Markov chain Monte Carlo algorithms, like
Riemmanian manifold Hamilton Monte Carlo, which can exploit the
geometry of the Grassmann manifold, may be useful here \citep{Byrne2013,
  Girolami2011}.  It may also be possible, though computationally
intensive, to jointly estimate the rank of the shared subspace, using
for instance, a reversible-jump MCMC algorithm.

Fundamentally, our approach is quite general and can be integrated
with other existing approaches for hierarchical multi-group covariance
estimation.  In particular, we can incorporate additional shrinkage on
the projected covariance matrices $\psi_k$.  As in \citet{Hoff2009} we
can employ non-uniform Bingham prior distributions for the
eigenvectors of $\psi_k$ or model $\psi_k$ as a function of continuous
covariates as in \citet{Yin2010, Hoff2011}.  Alternatively, we can
summarize the estimated covariance matrices by thresholding entries of
the precision matrix, $\psi_k^{-1}$ to visualize differences between
groups using a graphical model \citet{Meinshausen2006}.  The specifics
of the problem at hand should dictate which shrinkage models are
appropriate, but the shared subspace assumption should be useful in a
wide range of analyses, especially when the number of features is very
large.

\bibliographystyle{plainnat}
\bibliography{refs.bib}

%\section{Appendix}

\end{document}