### 2016-07-06 paper/mscov.tex
Minor edits from proofread

\documentclass[12pt]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{textcomp}
\usepackage{longtable}
\usepackage{multirow, booktabs}
\usepackage{natbib}
\usepackage{url}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{chemarrow}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{tikz}
\usepackage[margin=.75in]{geometry}

\renewcommand{\baselinestretch}{1.5}
\newcommand{\bl}[1]{{\mathbf #1}}
\newcommand{\bs}[1]{{\boldsymbol #1}}
\newcommand{\tr}{\text{tr}}
\newcommand{\etr}{\text{etr}}
\newcommand{\Exp}[1]{{\text{E}}[ \ensuremath{ #1 } ]  }

\begin{document}

\title{Shared Subspace Models for Multi-Group Covariance Estimation
%
  \protect\thanks{Alexander M. Franks is a Moore/Sloan Data Science
    and WRF Innovation in Data Science Postdoctoral Fellow and a
    graduate of the Department of Statistics at Harvard University
    (\href{mailto:amfranks@uw.edu}{amfranks@uw.edu}).  Peter D. Hoff
    is a Professor in the Departments of Statistics, Biostatistics and
    the Center for Statistics and the Social Sciences at the
    University of Washington
    (\href{mailto:pdhoff@uw.edu}{pdfhoff@uw.edu}).
%
This work was partially supported 
 by the Washington Research Foundation Fund for Innovation in Data-Intensive Discovery and the Moore/Sloan Data Science Environments Project at the University of Washington.
%
 The authors are grateful to Dr.\ Daniel Promislow (Department of
 Pathology, University of Washington) and Dr.\ Jessica Hoffman
 (University of Alabama at Birmingham) for sharing data and ideas
 which contributed to framing of this paper.}}
\author{Alexander Franks and Peter Hoff} \date{\today}
\maketitle

\begin{abstract}

% From Hoff Although the covariance matrices corresponding to different
% populations are unlikely to be exactly equal they can still exhibit a
% high degree of similarity. For example, some pairs of variables may be
% positively correlated across most groups, whereas the correlation
% between other pairs may be consistently negative. In such cases much
% of the similarity across covariance matrices can be described by
% similarities in their principal axes, which are the axes that are
% defined by the eigenvectors of the covariance matrices. Estimating the
% degree of across-popu- lation eigenvector heterogeneity can be helpful
% for a variety of estimation tasks.

  We develop a model-based method for evaluating heterogeneity among
  several $p\times p$ covariance matrices in the big $p$, small $n$ setting.
  This is done by assuming a spiked covariance model for each group
  and sharing information about the space spanned by the group-level
  eigenvectors.  We use an empirical Bayes method to identify a low-dimensional
  subspace which explains variation across all groups and use an MCMC
  algorithm to estimate the posterior uncertainty of eigenvectors and
  eigenvalues on this subspace.  The implementation and utility of our
  model is illustrated with analyses of high-dimensional multivariate
  gene expression and metabolomics data.
\vfill
\noindent\textbf{Keywords:} covariance estimation; spiked covariance
model; Stiefel manifold; Grassmann manifold; big p, small n;
high-dimensional data; empirical Bayes.
\end{abstract}

\section{Introduction}

% For example, it has long been known that the sample covariance matrix
% is a poor estimator of the population covariance matrix, particularly
% when the sample size is comparable to or smaller than the number of
% features ($n \leq p$) \citep{Dempster1969, Stein1975}.  A host of
% estimators have been devised to address this fact. These primarily
% include shrinkage approaches \citep{Schafer2005, Ledoit2011} and low
% rank factor models \citep{Mardia1980, Ullman2003, Fan2008}.  In many
% cases, a covariance matrix is highly structured and thus can be
% sparsely parameterized even if the matrix itself is not low rank
% \citep{Burg1982, Williams1993} .  Along these lines, there has been a
% recent interest in sparse Gaussian graphical model, in which the there
% are assumed to be many zeros in the inverse covariance matrix
% \citep{Meinshausen2006, Friedman2008}.

Multivariate data is often partitioned into
groups, each of which represent samples from populations with
distinct, but possibly related, distributions.  Although historically the
primary focus has been on identifying mean-level differences between
populations, there has been a growing need to identify differences in
population covariances as well.  For instance, in case-control studies, mean level effects may be small relative to subject
variability; distributional differences between groups may still be
evident as differences in the covariances between features.  Even when
mean level differences are detectable, better estimates of the
covariability of features across groups may lead to an improved
understanding of the mechanisms underlying these apparent mean level
differences.  Further, accurate covariance estimation is an essential
part of many prediction tasks (e.g. quadratic discriminant analysis).
Thus, evaluating heterogeneity between covariance
matrices can be an important complement to more traditional analyses
for estimating differences in means across
groups.

To address this need, we develop a novel method for
group-covariance estimation.  In our method, we exploit the fact that
in many natural systems, high dimensional data is often very
structured and thus can be best understood on a lower dimensional
subspace. For example, with gene expression data, the effective
dimensionality is thought to scale with the number of gene regulatory
modules, not the number of genes themselves \citep{Heimberg2016}.  As
such, differences in gene expression across groups should be expressed
in terms of differences between these regulatory modules rather than
strict differences between expression levels.  Such differences can be
examined on a subspace that reflects the correlations resulting from
these modules.  In contrast to most existing approaches for group
covariance estimation, our approach is to directly infer such
subspaces from groups of related data.

% Existing approaches incorporate a variety of
% different structural assumptions about the relationship between
% multiple groups of covariance matrices.

Some of the earliest approaches for multi-group covariance estimation
focus on estimation in terms of spectral decompositions.
\cite{Flury1987} developed estimation and testing procedures for the
``common principal components'' model, in which a set of covariance
matrices were assumed to share the same eigenvectors.
\citet{Schott1991, Schott1999} considered cases in which only certain
eigenvectors are shared across populations, and \citet{Boik2002}
described an even more general model in which eigenvectors can be
shared between some or all of the groups.  More recently,
\citet{Hoff2009}, noting that eigenvectors are unlikely to be shared
exactly between groups, introduced a hierarchical model for
eigenvector shrinkage based on the matrix Bingham distribution.  There
has also been a significant interest in estimating covariance matrices
using Gaussian graphical models. For Gaussian graphical models, zeros
in the precision matrix correspond to conditional independencies
between pairs of features given the remaining features
\citep{Meinshausen2006}.  \citet{Witten2014} extended existing work in
this area to the multi-group setting, by pooling information about the
pattern of zeros across precision matrices.

% When the
% data can be grouped according to a set of continuous covariates
% continuously parameterized covariance models can be appropriate
% \citep{Chiu1996, Yin2010, Hoff2011}.

Another popular method for modeling relationships between
high-dimensional multivariate data is partial least squares regression
(PLS) \citep{Wold2001}. This approach, which is a special case of a bilinear factor
model, involves projecting the data onto a lower dimensional space which
maximizes the similarity of the two groups.  This technique does not
require the data from each group to share the same feature set.  A
common variant for prediction, partial least squares discriminant
analysis (PLS-DA) is especially common in chemometrics and
bioinformatics \citep{Barker2003} .  Although closely related to the
approaches we will consider here, the primarily goal of of PLS-based
models is to create regression or discrimination models, not to
explicitly infer covariance matrices from multiple groups of data.
Nevertheless, the basic idea that data can often be well represented
on a low dimensional space is an appealing one that we leverage.

In this paper we propose a multi-group covariance estimation model by
sharing information about the subspace spanned by group-level
eigenvectors.  The shared subspace assumption can be used to improve
estimates and facilitate the interpretation of differences between
covariance matrices across groups.  For each group, we assume ``the
spiked covariance model'' (also known as the ``partial isotropy
model''), a well studied variant of the factor model
\citep{Mardia1980, Johnstone2001}.  In Section \ref{sec:shared} we
briefly review the behavior of spiked covariance models for estimating
a single covariance matrix and then introduce our extension to the
multi-group setting.  In Section \ref{sec:inference} we describe an
empirical Bayes algorithm for inferring the shared subspace and
estimating the posterior distribution of the covariance matrices of
the data projected onto this subspace.

In Section \ref{sec:simulation} we investigate the behavior of this
class of models in simulation and demonstrate how the shared subspace
assumption is widely applicable, even when there is little similarity
in the covariance matrices across groups.  In particular, shared
subspace models include those in which there is no information shared
across groups because in the worst case we can assume the subspace
spanned by the group-level eigenvectors is simply the full
$p$-dimensional space.  In Section \ref{sec:app} we demonstrate the
utility of a shared subspace model in an analysis of gene expression
data from juvenile leukemia patients .  Despite the large feature size
($p > 3000$) and small sample size ($n < 100$ per group), we identify
interpretable similarities and differences in gene covariances on a
low dimensional subspace.  Finally, we conclude with a brief
exploration of heterogeneity in metabolomic data from a study on aging
in the fly species \textit{Drosophila melanogaster}.

\section{A Shared Subspace Spiked Covariance Model}
\label{sec:shared}

%%\paragraph{The Spiked Covariance Model:}

Suppose $S$ has a possibly degenerate Wishart$(\Sigma,n$)
distribution with density given by
\begin{equation} 
p(S | \Sigma, n) \propto l(\Sigma: S) =  |\Sigma|^{-n/2} \text{etr}( - \Sigma^{-1}  S/2 ) ,  
\label{eqn:lik}
\end{equation}
%
\noindent where $\Sigma \in \mathcal S_p^+$ and $n$ may be less than $p$.  Such
a likelihood results from $S$ being, for example, a residual sum
of squares matrix from a multivariate regression analysis. In this
case, $n$ is the number of independent observations minus the rank of
the design matrix.  The spiked principle components model (spiked PCA)
studied by \citet{Johnstone2001} and others assumes that
\begin{equation} 
\Sigma = \sigma^2 ( U  \Lambda  U^T  + I )
\label{eqn:spiked}
\end{equation}
%
\noindent where for $r \ll p$, $\Lambda$ is an $r\times r$ diagonal
matrix and $U \in \mathcal V_{p,r}$.  $\mathcal V_{p,r}$ is the
Stiefel manifold, which consists of the set of $p \times r$
orthonormal matrices in $\mathbb{R}^p$, so that $U^TU =
I$.  The spiked covariance formulation is appealing because it explicitly
partitions the covariance matrix into a tractable low rank ``signal''
and isotropic ``noise''.

%  For such a covariance matrix, we have
% $\Sigma^{-1} = (\bl U \Lambda \bl U^T + \bl I )^{-1} = \bl I - \bl U
% \Omega \bl U^T $
% where $\Omega = \text{diag}(\omega_1,\ldots, \omega_r)$ with
% $\omega_s = \lambda_j/(\lambda_s+1)$.  The likelihood (\ref{eqn:lik})
% in terms of $(\sigma^2,\bl U, \Omega)$ is then
% \begin{equation}
% L(\sigma^2,\bl U , \Omega : \bl Y)  = 
% (\sigma^2)^{-np/2} \etr(-\bl S/[2\sigma^2]) \times 
%    \etr( \Omega \bl U^T [\bl S/(2\sigma^2)] \bl U ) \times  
%  \left\{ \prod_{s=1}^r (1-\omega_s)^{n/2} \right \} .  
% \label{eqn:rplik}
% \end{equation}

Classical results for parametric models (e.g., \citet{Schwartz1965})
imply that asymptotically in $n$ for fixed $p$, an estimator will be
consistent for a spiked population covariance as long as the assumed
number of spikes (eigenvalues larger than $\sigma^2$) is greater than
or equal to the true number.  However, when $p$ is large relative to
$n$, as is the case for the examples considered here, things are more
difficult.  Under the spiked covariance model, it has been shown that
if $p/n \rightarrow \alpha >0$ as $n\rightarrow \infty$, the $k$th
largest eigenvalue of $S/(n\sigma^2)$ will converge to an upwardly
biased version of $\lambda_{k}+1$, if $\lambda_k$ is greater than
$\sqrt{\alpha}$ \citep{Baik2006, Paul2007} .  This has led several
authors to suggest estimating $\Sigma$ via shrinkage of the
eigenvalues of the sample covariance matrix. In particular, in the
setting that $\sigma^2$ is known, \citet{Donoho2013} propose
estimating all eigenvalues whose sample estimates are smaller than
$\sigma^2(1+\sqrt{\alpha})^2$ by $\sigma^2$, and shrinking the larger
eigenvalues in a way that depends on the particular loss function
being used.  These shrinkage functions are shown to be asymptotically
optimal in the $p/n\rightarrow \alpha$ setting.

Note that covariance estimators of this form are are equivariant with
respect to rotations and scale changes.  The situation should be
different, however, when we are interested in estimating multiple
covariance matrices from distinct, but related groups.  Here,
group-level equivariance to rotations is an unreasonable assumption;
both eigenvalue \emph{and} eigenvector shrinkage can play an important
role in improving covariance estimates.  Consider multi-group
covariance estimation based on $K$ matrices, $Y_1, ..., Y_K$, where
$Y_k$ is assumed to be an $n_k$ by $p$ matrix of mean-zero normal
data, typically with $n_k \ll p$.  Then, $Y_k^TY_k = S_k$ has a
(degenerate) Wishart distribution as in Equation \ref{eqn:lik}.  To
improve estimation, we seek estimators of each covariance matrix,
$\hat{\Sigma}_k$, that may depend on data from all groups.  To this
end, we extend the spiked covariance model to grouped data, by
assuming that the anisotropic variability from each group occurs on a
common low dimensional subspace.  Specifically, we posit that the
covariance matrix for each group can be written as
\begin{equation}
\Sigma_k = V\Psi_kV^T + \sigma^2_kI,
\label{eqn:sspsi}
\end{equation}

% As demonstrated by Gavish et al \citep{Gavish2014}, even when the
% true rank is large, if blah it is still asymptotically preferable to fit a lower rank
% blah .  it is always better than hard thresholding at any other value, and is always better than ideal truncated singular value decomposition (TSVD)

% Many eigenvalue shrinkage estimators can be viewed in the Bayesian
% context as derivative of a Bayes estimator under a particular prior.

%\paragraph{The Shared Subspace Model:}

\begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.25\textwidth}
        \includegraphics[width=\textwidth]{Figs/3dplot}
        \caption{Projection in $\mathcal{R}^3$}
        \label{fig:3dplot}
    \end{subfigure}
\quad
 %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.25\textwidth}
        \includegraphics[width=\textwidth]{Figs/2d-scatter}
        \caption{$Y_kV$}
        \label{fig:2dscatter}
    \end{subfigure}
\quad 
    \begin{subfigure}[b]{0.25\textwidth}
        \includegraphics[width=\textwidth]{Figs/2d-scatter-orth}
        \caption{$Y_kV_{\perp}$}
        \label{fig:2dscatterorth}
    \end{subfigure}
    \caption{ Two groups of four-dimensional data (red and blue)
      projected into different subspaces.  a) To visualize $Y_k$ we
      can project the data into $R^3$.  In this illustration, the
      distributional differences between the groups are confined to a
      two-dimensional shared subspace ($VV^T$, grey plane).  b) The
      data projected onto the two-dimensional shared subspace,
      $Y_kV$, have covariances $\Psi_k$ that differ between
      groups. c) The orthogonal projection, $Y_kV_{\perp}$
      has isotropic covariance, $\sigma_k^2I$, for all groups.  }
\label{fig:shared}
\end{figure}
% For multi-group covariance estimation, we assume that we have $K$
% matrices, $Y_1, ... Y_K$, where $Y_k$ consists of $n_k$ independent
% rows of mean-zero normal data of dimension $p$, typically with
% $n_k \ll p$.

%
\noindent where, the columns of $V \in \mathcal{V}_{p, s}$, with
$s \ll p$, determine a subspace shared by all groups.  Throughout this
paper we will denote to the shared subspace as
$VV^T \in \mathcal G_{p,s}$, where $\mathcal G_{p,s}$, known as the
Grassmann manifold, consists of all $s$-dimensional linear subspaces
of $\mathbb{R}^p$ \citep{Chikuse2012}.  Although $V$ is only
identifiable up to right rotations, the matrix $VV^T$, which defines
the plane of variation shared by all groups, is identifiable.  Later,
to emphasize the connection to the spiked PCA model
(\ref{eqn:spiked}), we will write $\Psi_k$ in terms of its
eigendecomposition, $\Psi_k = O_k\Lambda_kO_k$, where $O_k$ are
eigenvectors and $\Lambda_k$ are the eigenvalues of $\Psi_k$ (see
Section \ref{sec:bayes}).

Note that for a shared subspace model,
$V^T\Sigma_kV = \Psi_k + \sigma_k^2I$ is an anisotropic
$s$-dimensional covariance matrix for the projected data, $Y_kV$.  In
contrast, the data projected onto the orthogonal space,
$Y_kV_{\perp}$, is isotropic for all groups.  In Figure
\ref{fig:shared} we provide a simple illustration using simulated
$4$-dimensional data from two groups.  In this example, the
differences in distribution between the groups of data can be
expressed on a two dimensional subspace spanned by the columns of
$V \in \mathcal{V}_{4, 2}$.  Differences in the correlations between
the two groups manifest themselves on this shared subspace, whereas
only the magnitude of the isotropic variability can differ between
groups on the orthogonal space.  Thus, a shared subspace model
can be viewed as a covariance partition model, where one partition
includes the anisotropic variability from all groups and the other
partition is constrained to the isotropic variability from each group.
This isotropic variability is often characterized as measurement
noise.

%Such a modelis reasonable for many applications involving groups of
%highly structure data on the same set of $p$ features.  % For instance,
% in biological applications, only certain subsets of molecules can
% physically interact which constrains the space of possible correlation
% patterns.  Our model exploits this notion, by estimating differences
% between groups on a common lower dimensional subpace.

\section{Empirical Bayes Inference}
\label{sec:inference}

In this section we outline an empirical Bayesian approach for
estimating a low-dimensional shared subspace and the covariance
matrices of the data projected onto this space. As we highlight in
Section \ref{sec:simulation}, if the spiked covariance model holds for
each group individually, then the shared subspace assumption also
holds because at worst we can take the shared subspace to be the full
$p$-dimensional space. In practice, we can usually identify a shared
subspace of dimension $s \ll p$ that preserves most of the variation
in the data.  Our primary objective is to identify the ``best'' shared
subspace of fixed dimension $s < p$.  In Section \ref{sec:em} we
describe an expectation maximization algorithm for estimating the
maximum marginal likelihood of the shared subspace, $VV^T$.  This
approach is computationally tractable for high-dimensional datasets.
Given an inferred subspace, we then seek estimators for the covariance
matrices of the data projected onto this space.  Because seemingly
large differences in the point estimates of covariance matrices across
groups may not actually reflect statistically significant differences,
in Section \ref{sec:bayes} we also describe a Gibbs sampler that can
be used to generate estimates of the projected covariance matrices,
$\Psi_k$, and and their associated uncertainty.  Later, we discuss
strategies for infering an appropriate value for $s$ and explore how
shared subspace models can be used for exploratory data analysis by
visualizing covariance heterogeneity on two or three dimensional
subspaces.

% However, in the $n \ll p$ setting, seemingly large
% differences in the point estimates of the covariance matrices across
% groups may not actually reflect statistically significant differences.
% Thus, in Section \ref{sec:bayes} we describe a Gibbs sampler that can
% be used to infer both point estimates and the associated uncertainty
% of the projected covariance matrices, $\Psi_k$.

%TODO: emphasize 1) justification 2) objective of \emph{identifying} this
%subspace when p much greater than n.

\subsection{Estimating the  Shared Subspace}
\label{sec:em}

In this section we describe a maximum marginal likelihood procedure for
estimating the shared subspace, $VV^T$, based on the
expectation-maximization (EM) algorithm.  The full likelihood
for the shared subspace model can be written as

\begin{align}
\nonumber p(S_1, ... S_k | \Sigma_k,n_k) &\propto \prod_{k=1}^K |\Sigma_k|^{-n_k/2}\etr(-\Sigma_k^{-1}S_k/2)  \\
\nonumber &\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}\etr(-(\sigma_k^2(V\Psi_kV^T +
  I))^{-1}S_k/2) \\
\nonumber &\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}\etr(-\left[V(\Psi_k +
  I)^{-1}/\sigma_k^2 V^T + (I-VV^T)/\sigma^2_k\right]S_k/2)
  \\
&\propto \prod_{k=1}^K  (\sigma_k^2)^{-n_k(p-s)/2}|M_k|^{-n_k/2}\etr(-\left[VM_k^{-1}V^T + \frac{1}{\sigma^2_k} (I-VV^T)\right]S_k/2) ,
\end{align}
%
\noindent where we define $M_k = \sigma_k^2(\Psi_k + I)$.  The log-likelihood in
$V$ (up to an additive constant) is
%
\begin{align}
\nonumber l(V) &= \sum_k \tr\left(-VM_k^{-1}V^T +
       VV^T/\sigma^2_k\right)S_k/2)\\
&= \sum_k \frac{1}{2}\tr\left(-M_k^{-1}V^T S_kV\right) + \frac{1}{2\sigma_k^2}tr\left(VV^T S_k\right).
\end{align}

% \begin{align}
% p(V|...) &\propto \etr(\sum_{k=1}^K[O_k\Omega_kO^T_k]V^TS_k/(2\sigma^2_k)]V)\\
% \end{align}

We maximize the marginal likelihood of $V$ with an EM algorithm, where
$M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as the considered the
``missing'' parameters to be imputed.  We assume independent Jeffreys'
prior distributions for both $\sigma_k^2$ and $M_k$.  Jeffreys' prior
for these quantities corresponds to
$p(\sigma_k^2) \propto 1/\sigma_k^2$ and
$p(M_k) \propto |M_k|^{-(s+1)/2}$.  From the likelihood it can easily
be shown that the conditional posterior for $M_k$ is
%
$$p(M_k | V) \propto |M_k|^{(n_k + s + 1)/2}\etr(-1/2(M_k^{-1}V^TS_kV)) $$
%
\noindent which is an inverse-Wishart($V^TS_kV$, $n_k$) distribution.  The
conditional posterior distribution of $\sigma_k^2$ is simply
%
$$p\left(\sigma^2 | V\right) \propto (\sigma_k^2)^{-n_k(p-s)/2-1}\etr\left(-\frac{1}{2\sigma^2_k} (I-VV^T)S_k\right)  $$
%
\noindent which is an inverse-gamma($n_k(p-s)/2$,
$\frac{1}{2}\text{tr}[(I-VV^T)S_k]$) distribution.  Based on these
results, we complete the following steps for each iteration, $t$, of
the EM algorithm until convergence:

\begin{enumerate}
\item For each $k$, compute relevant conditional expectations:
\begin{enumerate}
\item $E[M_k^{-1} | V_{(t-1)}] = n_k(V_{(t-1)}^T S_kV_{(t-1)})^{-1}$
\item $E[\frac{1}{\sigma_k^2}|V_{(t-1)}] = \frac{n_k(p-s)}{\text{tr}[(I-V_{(t-1)}V_{(t-1)}^T)S_k]}$
\end{enumerate}
\item Compute $V_{t}$ = $\underset{V}{\text{arg } \text{max}}  \sum_k \text{tr}\left(-VE[M^{-1}|V_{(t-1)}]^T +
       E[\frac{1}{2\sigma_k^2}|V_{(t-1)}]VV^T\right)S_k/2)$ 
\end{enumerate}

For the second step (``M-step''), we use a numerical optimization algorithm based on
the Cayley transform to preserve the orthogonality constraints in $V$
\citep{Wen2013}.  Importantly, the complexity of this algorithm is
dominated by the dimension of the shared subspace not the number of
features.  Thus, our approach is computationally efficient for
relatively small values of $s$, even when $p$ is large.

\paragraph{Evaluating Goodness of Fit:}

If $V$ is a basis for a shared subspace, then for each group, most
of the non-isotropic variation in $Y_k$ should be preserved when
projecting the data to this space.  To characterize the extent to
which this is true for different groups, we propose a simple estimator
for the proportion of ``signal'' variance that lies on a given
subspace.  Specifically, we use the following statistic for the ratio
of the sum of the first $s$ eigenvalues of
$V^T \Sigma_k V$ to the sum of the first $s$ eigenvalues
of $\Sigma_k$:
%
\begin{equation}
\gamma(Y_k: V, \sigma_k^2) = \frac{||Y_kV||_F/n_k}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}
  ||Y_k\widetilde{V}||_F/n_k - \sigma_k^2ps/n_k}
\label{eqn:ratio}
\end{equation}
%
\noindent where $||\cdot||_F$ is the Frobenius norm. Our motivation for
this statistic is the fact that if
$\Sigma_k = V\Psi_kV^T + \sigma_k^2I$ then
$\gamma(Y_k: V, \sigma_k^2) \approx 1$.  We establish
this in an asymptotic regime where $s$ is fixed, $p$,
$n_k \rightarrow \infty$ and $p/n_k = \alpha_k$ constant.

First, since $s$ is fixed and $n_k$ is growing, the numerator,
$||Y_kV||_F/n_k$, is a consistent estimator for the sum of the
eigenvalues of $V^T\Sigma_kV$.  The denominator,
$\underset{\widetilde{V} \in \mathcal{V}_{p,
    s}}{\text{max}}||Y_k\widetilde{V}||_F/n_k$
is equivalent to the sum of the first $s$ eigenvalues of the sample
covariance matrix $S_k/n_k$.  With $\hat{\lambda}^{(k)}_i$ the
$i$-th eigenvalue of $S_k/n_k$, \citet{Baik2006} demonstrated that
asymptotically as $p, n_k \rightarrow \infty$ with $p/n_k = \alpha_k$
fixed
%
\begin{eqnarray}
\nonumber \hat{\lambda}^{(k)}_i &\rightarrow& \lambda^{(k)}_i\left(1 +
                                    \frac{\sigma_k^2\alpha_k}{\lambda^{(k)}_i
                                    - 1}\right)\\
& \approx& \lambda^{(k)} + \sigma^2_k\frac{p}{n_k}
\label{eqn:lamAsymp}
\end{eqnarray}
%
\noindent where the approximation in the second line is due to the
fact that $\lambda_k/(\lambda_k - 1) \approx 1$ for $\lambda_k$ large.
Thus,
$\sum_i^s \hat{\lambda}^{(k)}_i \approx \sum_i^s \lambda^{(k)}_i +
\sigma_k^2ps/n_k$
and
$\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}
||Y_k\widetilde{V}||_F/n_k - \hat{\sigma}_k^2ps/n_k$
is a reasonable approximation to the sum of the first $s$ eigenvalues
of $\Sigma_k$.  Alternatively, when $\lambda_k$ is small, a better
approximation may be obtained by solving the quadratic equation
implied by the first line of Equation \ref{eqn:lamAsymp}.

As a consequence, the proposed statistic will be close to one for all
groups when $VV^T$ is a shared subspace for the data and smaller if
not.  The metric provides a useful indicator for which groups can be
reasonably compared on a given subspace and which groups cannot.  In
practice, we estimate a shared subspace, $\hat{V}$, and the isotropic
variances, $\hat{\sigma}_k^2$, using EM and compute the plug-in
estimate $\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2)$.  When this
statistic is small for some groups, it may suggest that the rank, $s$,
of the inferred subspace needs to be larger to capture the variation
in all groups. We investigate this in Section \ref{sec:simulation}, by
computing the goodness of fit statistic for inferred subspaces of
different dimensions on a single dataset. In Section \ref{sec:app}, we
compute the estimates for subspaces inferred with real biological
data.

%% Problem here! Can't compare different sample sizes??

\subsection{Inference for Projected Covariance Matrices}
\label{sec:bayes}

The EM algorithm presented in the previous section yields point
estimates for $VV^T$ and $\Psi_k$, but does not lead to natural
uncertainty quantification for these estimates.  In this section, we
assume that the subspace $VV^T$ is fixed and known and demonstrate how
we can estimate the posterior distribution for $\Psi_k$.   Note that when
the subspace is known, the posterior distribution of $\Sigma_k$ is
conditionally independent from the other groups, so that we can
independently estimate the conditional posterior distributions for each
group.

%  In this paper, we build on recent interest in the spiked
% covariance model \citep{Donoho2013, Paul2007} and develop tractable
% models for the eigendecomposition of $\Psi_k$.  Such models are
% especially appealing because it is simple to summarize and compare
% $\Psi_k$ across groups in terms their eigenstructure (See Section
% \ref{sec:simulation} and \ref{sec:app}).

There are many different ways in which we could choose to parameterize
$\Psi_k$.  Building on recent interest in the spiked covariance model
\citep{Donoho2013, Paul2007} we propose a tractable MCMC algorithm by
specifying priors on the eigenvalues and eigenvectors of $\Psi_k$.  By
modeling the eigenstructure, we can now view each covariance
$\Sigma_k$ in terms of the original spiked principle components model.  Equation \ref{eqn:sspsi}, written as a function of
$V$, becomes
%
\begin{align}
\nonumber \Psi_k &= O_k\Lambda_KO_k^T\\
\Sigma_k &= V\Psi_kV^T + \sigma^2_kI.
\label{eqn:ss}
\end{align}
%
\noindent Here, we allow $\Psi_k$ to be of rank $r \leq s$ dimensional
covariance matrix on the $s$-dimensional subspace.  Thus, $\Lambda_k$
is an $r \times r$ diagonal matrix of eigenvalues, and
$O_k \in \mathcal{V}_{s,r}$ is the matrix of eigenvectors of $\Psi_k$.
For any individual group, this corresponds to the original spiked PCA
model (Equation \ref{eqn:spiked}) with
$U_k = VO_k \in \mathcal{V}_{p, r}$.  Differentiating the ranks $r$
and $s$ is helpful because it enables us to independently specify a
subspace common to all groups and the possibly lower rank features on
this space that are specific to individual groups.  Although our model
is most useful when the covariance matrices are related across groups,
we can also use this formulation to specify models for multiple
unrelated spiked covariance models.  We explore this in detail in
Section \ref{sec:simulation}.  In Section \ref{sec:app} we introduce a
shared subspace model with additional structure on the eigenvectors
and eigenvalues of $\Psi_k$ to facilitate interpretation of covariance
heterogeneity on a two-dimensional subspace.

The likelihood for $\Sigma_k$ given the sufficient statistic
$S_k = Y_kY_k^T$ is given in Equation \ref{eqn:lik}.  For the
spiked PCA formulation, we must rewrite this likelihood in terms of $V$, $O_k$,
$\Lambda_k$ and $\sigma_k^2$.  First note that by the Woodbury matrix
identity
 %
\begin{align}
\nonumber \Sigma^{-1}_k &=  (\sigma_k^2(U_k\Lambda_kU_k^T+I))^{-1}\\
\nonumber &= \frac{1}{\sigma_k^2}(U_k\Lambda_kU_k^T+I)^{-1}\\
&= \frac{1}{\sigma_k^2}(I-U_k\Omega_kU_k^T),
\end{align}
%
\noindent where the diagonal matrix $\Omega = \Lambda(I+\Lambda)^{-1}$, e.g. $\omega_i = \frac{\lambda_i}{\lambda_{i}+1}$.  Further, 
%
\begin{align}
\nonumber |\Sigma_k| &= (\sigma_k^2)^{p}|U_k\Lambda_kU_k^T+I|\\
\nonumber &= (\sigma_k^2)^{p}|\Lambda_k+I| \\
\nonumber &= (\sigma_k^2)^{p}\prod_{i=1}^r(\lambda_i+1)\\
&= (\sigma_k^2)^{p}\prod_{i=1}^r(1-\omega_i),
\end{align}
%
\noindent where the second line is due to Sylvester's determinant
theorem.  Now, the likelihood of $V$, $O_k$, $\Lambda_k$ and
$\sigma_k^2$ is available from Equation \ref{eqn:lik} by substituting
the appropriate quantities for $\Sigma^{-1}_k$ and $|\Sigma_k|$ and
replacing $U_k$ with $VO_k$:
%
\begin{equation}
 L(\sigma_k^2, V , O_k \Omega_k : Y_k) \propto
    (\sigma_k^2)^{-n_kp/2}\etr(-\frac{1}{2\sigma_k^2}S_k)\left(\prod_{i=1}^r(1-\omega_{ki})
   \right) ^{n_k/2}
   \etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)S_k).
\label{eqn:sslik}
\end{equation}
%
\noindent We use conjugate and semi-conjugate priors for the parameters $O_k$,
$\sigma^2_k$ and $\Omega_k$ to facilitate inference via a Gibbs
sampling algorithm.  In the absence of specific prior information,
invariance considerations suggest the use of priors that lead to
equivariant estimators.  Below we describe our choices for the prior
distributions of each parameter and the resultant conditional posterior
distributions.

\paragraph{Conditional distribution of $\sigma_k^2$:}

From Equation \ref{eqn:sslik} it is clear that the the inverse-gamma
class of prior distributions is conjugate for $\sigma_k^2$.  We chose a
default prior distribution for $\sigma^2_k$ that is equivariant with
respect to scale changes.  Specifically, we use Jeffreys' prior, an
improper prior with density $p(\sigma^2_k) \propto 1/\sigma^2_k $.  Under
this prior, straightforward calculations show that the full
conditional distribution of $\sigma_k^2$ is
inverse-gamma$( n_k p/2 , \tr[S_k( I - U_k\Omega_k
U_k)/2])$, where $U_k = VO_k$.

\paragraph{Conditional distribution of $O_k$:} Given the likelihood
from Equation \ref{eqn:sslik}, it is easy to show that the class of
Bingham distributions are conjugate for $O_k$ \citep{Hoff2009,
  Hoff2012}.  Again, invariance considerations lead us to use a
rotationally invariant uniform probability measure on
$\mathcal V_{s,p}$.  Under this uniform prior, the full conditional
distribution of $O_k$ has a density proportional to the
likelihood, so
\begin{align}
\label{lik_vo}
 p(O_k | \sigma^2_k, U_k, \Omega_k) & \propto \etr(\Omega_kO^T_kV^T[S_k/(2\sigma^2_k)]VO_k).
\end{align}
%
\noindent This is a Bingham$(\Omega, V^T S_k V/(2\sigma^2))$
distribution on $\mathcal V_{s, r}$ \citep{Khatri1977}. A
Gibbs sampler to simulate from this distribution is given in
\citet{Hoff2012}.

Together, the prior for $\sigma_k^2$ and $O_k$ leads to conditional
(on $V$) Bayes estimators $\hat \Sigma(V^T S_k V)$ that are
equivariant with respect to scale changes and rotations on the
subspace spanned by $V$, so
that $\hat \Sigma(a W V^T S_k V W^T) = a W \hat\Sigma(V^T
S_k V)  W$
for all $a>0$ and $ W\in \mathcal O_{s}$ (assuming an invariant
loss function). Interestingly, if $\Omega_k$ were known (which it is
not), then for a given invariant loss function the Bayes estimator
under this prior minimizes the (frequentist) risk among all
equivariant estimators \citep{Eaton1989}.

\paragraph{Conditional distribution for $\Omega_k$:} Here we specify the
conditional distribution of the diagonal matrix
$\Omega_k = \Lambda_k(I+\Lambda_k)^{-1} = \text{diag}(\omega_{k1},
... \omega_{kr})$.
We consider a uniform(0,1) prior distribution for each element of
$\Omega$, or equivalently, an $F_{2,2}$ prior distribution for the
elements of $\Lambda$.  The full conditional distribution of an
element $\omega_i$ of $\Omega$ is proportional to the likelihood
function, so

\begin{align}
p(\omega_{ki}|V, O_k, S_k) &\propto_{\omega_{ki}}
  \left(\prod_{i=1}^r(1-\omega_{ki})^{n_k/2}  \right)
  \etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)S_k) \\
&  \propto  (1-\omega_{ki})^{n/2} e^{c_{ki} \omega_{ki}  n/2},    
\label{eqn:wpost}
\end{align}
%
\noindent where $c_{ki} = u_{ki}^T S_k u_{ki}/(n_k \sigma^2_k)$ and $ u_{ki}$ is
column $i$ of $U_k = VO_k$.  While not proportional to a density
belonging to a standard class of distributions, we can sample from the
corresponding univariate distribution numerically.  The behavior of
this distribution is straightforward to understand: if $c_{ki}\leq 1$, then
the the function has a maximum at $\omega_{ki} =0$, and decays
monotonically to zero as $\omega_{ki} \rightarrow 1$.  If $c_{ki}>1$ then the
function is uniquely maximized at $(c_{ki}-1)/c_{ki} \in (0,1)$.  To see why
this makes sense, note that the likelihood is maximized when the
columns of $ U_k$ are equal to the eigenvectors of $S_k$
corresponding to its top $r$ eigenvalues
\citep{Tipping1999}. At this value of $U_k$, $c_{ki}$ will then
equal one of the top $r$ eigenvalues of $ S_k/(n_k\sigma_k^2)$.  In the
case that $n_k\gg p$, we expect
$ S_k/(n_k\sigma_k^2)\approx \Sigma_k/\sigma_k^2$, the true (scaled)
population covariance, and so we expect $c_{ki}$ to be near one of the top
$r$ eigenvalues of $\Sigma_k/\sigma^2_k$, say $\lambda_{ki}+1$.  If indeed
$\Sigma_k$ has $r$ spikes, then $\lambda_{ki}>0$,
$c_{ki} \approx \lambda_{ki} +1 > 1$, and so the conditional mode of $w_{ki}$ is
approximately $(c_{ki}-1)/c_{ki} = \lambda_{ki}/(\lambda_{ki}+1)$, the correct value.
On the other hand, if we have assumed the existence of a spike when
there is none, then $\lambda_{ki}=0$, $c_{ki}\approx 1$ and the Bayes estimate
of $w_{ki}$ will be shrunk towards zero, as it should be.

% \paragraph{Conditional distribution of $V$}
% Need to finish / should we include?

% Although we prefer EM... it is also posible to estimate $V$ using
% Bayesian inference.  We assume a uniform prior on $V \in \mathcal{V}_{p, s}$ so that the
% conditional posterior of $V$ is proportional to the likelihood:

% \begin{align}
% p(V|\bl O, \bl S, \bl \sigma^2, \bl \Omega) &\propto \etr(\sum_{k=1}^K[S_k/(2\sigma^2_k)]V[O_k\Omega_kO^T_k]V^T)\\
% \end{align}

% Although close in appearance, this conditional distribution is not a matrix
% Bingham-von Mises-Fisher distribution.  However, the conditional distribution
% of a column of $V$ given all other columns is a vector Bingham-von
% Mises-Fisher distribution.  To see this, let
% $A_k = S_k/(2\sigma^2_k)$ and $B_k = O_k\Omega_kO^T_k$.  We can write
% the conditional distribution of $V$ given $A_k$ and $B_k$ as
% \begin{align}
% P(V|A_k,B_k) &\propto \etr(\sum_{k=1}^KA_kVB_kV^T)\\
% &= \etr(\sum_{k=1}^K\sum_{i,j}^sA_kv_iv_j^Tb_{ijk})\\
% &= \etr(\sum_{i,j}^s \sum_{k=1}^K v_j^T[b_{ijk}A_k]v_i)
% \end{align}

% From here, we specify the conditional distribution of a single column of $V$ given $A_k$, $B_k$ and all other columns of V:

% \begin{align}
% P(v_i|A_k,B_kV_{-i})  &\propto \etr(2(\sum_{j\neq i}^s v_j^T[\sum_{k=1}^K b_{ijk}A_k])v_i+v_i^T[\sum_{k=1}^K b_{ijk}A_k]v_i))\\
% &= \etr(Cv_i+v_i^TMv_i)
% \label{eqn:vbmf}
% \end{align}

% This is a vector Bingham-von Mises-Fisher Distribution, BMF(C, M)
% with $C = 2(\sum_{j\neq i}^s v_j^T[\sum_{k=1}^K b_{ijk}A_k])$ and $M=
% \sum_{k=1}^K b_{ijk}A_k$.  Following \citet{Hoff2012},  we can derive
% a Gibbs sampler for $V$ based on conditional draws of the columns of
% $V$.

\section{Simulation Study}
\label{sec:simulation}

%\subsection{Validation of Inference}

We start with an example demonstrating how a shared subspace
model can be used to identify statistically significant differences between
covariance matrices on a low dimensional subspace. Here, we simulate
$K=5$ groups of data from the shared subspace spiked covariance
model with $p=200$, $s=r=2$, $\sigma_k^2=1$, and
$n_k=50$.  We fix the first eigenvalue of $\Psi_k$ from each group to
$\lambda_1=100$ and vary $\lambda_2$.  For this two dimensional shared subspace
model we summarize $\Psi_k$ in terms of its eigendecomposition by
computing posterior distributions for the log eigenvalue ratio,
$\text{log}(\frac{\lambda_1}{\lambda_2})$, with $\lambda_1 > \lambda_2$, and the
angle of the first eigenvector on this subspace,
$\text{arctan}(\frac{O_{12}}{O_{21}})$, relative to the first column
of $V$.

Figure \ref{fig:simPosterior} depicts the 95\% posterior regions for
these quantities from a single simulation.  Dots correspond to the
true log ratios and orientations of $\hat{V}^T\Sigma_k\hat{V}$, where
$\hat{V}$ is the maximum marginal likelihood for $V$. To compute
the posterior regions, we iteratively remove posterior samples corresponding to the
vertices of the convex hull until only 95\% of the original samples
remain.  Non-overlapping posterior regions imply that differences in
the covariances are statistically significant between groups.  In this
example, the ratio of the eigenvalues of the true covariance matrices
were $10$ (black and red groups), $3$ (green and blue groups) and $1$
(cyan group).  Larger eigenvalue ratios correspond to more
correlated contours and a value of $1$ implies isotropic covariance.
Note that for the smaller eigenvalue ratio of $3$, there is larger
uncertainty about the orientation of the primary axis.  When the ratio
is one, as is the case for the cyan colored group, there is no
information about the orientation of the primary axis since the contours
are spherical.

\begin{figure}[t]
    \centering
%    \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=0.5\textwidth]{Figs/posteriorRegions-chull}
        \caption{95\% posterior regions}
        \label{fig:simPosterior}
%    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    % \begin{subfigure}[b]{0.4\textwidth}
    %     \includegraphics[width=\textwidth]{Figs/simRatio-05-19}
    %     \caption{Ratio}
    %     \label{fig:simRatio}
    % \end{subfigure}
    \caption{a) 95\% posterior regions for the log of the ratio of
      eigenvalues, $\text{log}(\frac{\lambda_1}{\lambda_2})$, of $\Psi_k$ and the
      orientation of the principle axis on the space spanned by $\hat{V}$
      cover the truth in this simulation.  Dots correspond to true data generating
      parameter values on $\hat{V}^T\Sigma_k\hat{V}$ .  Since $V$
      is only identifiable up to rotation, for this figure we find the Procrustes
      rotation that maximizes the similarity of $\hat{V}$ to the true
      data generating basis. True eigenvalue ratios were 10 (red and
      black), 3 (green and blue) and 1 (light blue).  True
      orientations were $\pi/4$ (black), $-\pi/4$ (red and green) and
      0 (blue and cyan). % b) Goodness of fit,
      % $\gamma(Y_k: \hat{V}, \hat{\sigma}^2_k$ (Equation \ref{eqn:ratio}).
      % As expected, for each group, the inferred subspace explains
      % practically all of the variance from the first two eigenvectors
      % of $\Sigma_k$. 
    }
\end{figure}

To demonstrate the overall validity of the shared subspace approach,
we compute the frequentist coverage of these 95\% bayesian credible
regions for the eigenvalue ratio and primary axis orientation using
one thousand simulations.  For the two groups with eigenvalue ratio
$\lambda_1/\lambda_2 = 3$ the frequentist coverage was close to
nominal at approximately 0.94.  For the groups with
$\lambda_1/\lambda_2 = 10$ the coverage was approximately 0.92.  We
did not evaluate the coverage for the group with
$\lambda_1/\lambda_2 = 1$ (cyan) since this value is on the edge of the
parameter space and is not be covered by the 95\% posterior regions
as constructed.  The slight under coverage for the other groups is likely due to the fact
that we infer $VV^T$ using maximum marginal likelihood, and thus
ignore the extra variability due to the uncertainty about the shared
subspace estimate.

\subsection{Rank Selection and Model Misspecification}

\begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{Figs/LossVsDimension}
        \caption{Stein's risk vs $\hat{s}$}
        \label{fig:sdimension}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{Figs/simRatio-s5}
        \caption{$\hat{s}$ = 5}
        \label{fig:ratio-s5}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{Figs/simRatio-s20}
        \caption{$\hat{s}$ = 20}
        \label{fig:ratio-s20}
    \end{subfigure}
    \caption{a) Stein's risk as a function of the
      shared subspace dimension.  Data from ten groups, with $U_k$
      generated uniformly on the Stiefel manifold
      $\mathcal{V}_{200, 2}$.  As $\hat{s} \rightarrow p$, the risk converges to the
      risk from independently estimated spiked covariance matrices
      (dashed blue line).  The data also fit a shared subspace model
      with $s=rK$.  If $VV^T = \text{span}(U_1, ..., U_k)$ were known
      exactly, shared subspace estimation yields lower risk than
      independent covariance estimation (dashed red line).  b) For a
      single simulated dataset, the goodness of fit statistic,
      $\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$, when the assumed
      shared subspace is dimension $\hat{s} = 5$.  c).  For the same
      dataset, goodness of fit when the assumed shared subspace is
      dimension $\hat{s} = 20$.  We can capture nearly all of the
      variability in each of the 10 groups using an $\hat{s}=rK=20$
      dimensional shared subspace. }
\label{fig:dimensionPlots}
\end{figure}

Naturally, shared subspace inference works well when the model is
correctly specified.  What happens when the model is not well
specified?  We explore this question in silico by simulating data from
different data generating models and evaluating the efficiency of
various covariance estimators.  In all of the following simulations we
evaluate covariance estimates using Stein's loss,
$L_S( \Sigma_k , \hat\Sigma_k) = \text{tr}( \Sigma_k^{-1} \hat
\Sigma_k ) - \log |\Sigma_k^{-1} \Sigma_k | - p$.
Since we compute multi-group estimates, we report the
average Stein's loss
$L(\Sigma_1, ..., \Sigma_K; \hat\Sigma_1, ..., \hat\Sigma_K ) =
\frac{1}{K} \sum_k L_S( \Sigma_k , \hat\Sigma_k)$.
Under Stein's loss, the Bayes estimator is the inverse of
the posterior mean of the precision matrix,
$\hat \Sigma_{k} = \Exp{ \Sigma_k^{-1} | S_k}^{-1}$ which we
estimate using MCMC samples.

We start by investigating the behavior of our model when we
underestimate the true dimension of the shared subspace.  In this
simulation, we generate $K=10$ groups of mean-zero normally
distributed data with $p=200$, $r=2$, $s=p$ and $\sigma_k^2=1$.  We
fix the eigenvalues of $\Psi_k$ to
($\lambda_1, \lambda_2) = (250, 25)$.  Although the signal variance
from each group individually is preserved on a two dimensional
subspace, these subspaces are not similar across groups since the
eigenvectors from each group are generated uniformly from the Stiefel
manifold, $U_k \in \mathcal{V}_{p, r}$.

We use this data to evaluate how well the shared subspace estimator
performs when we fit the data using a shared subspace model of
dimension $\hat{s} < s$.  In Figure \ref{fig:sdimension} we plot
Stein's risk as a function of $\hat{s}$, estimating the risk
empirically using ten independent simulations per value of $\hat{s}$.
The dashed blue line corresponds to Stein's risk for covariance
matrices estimated independently.  Independent covariance estimation
is equivalent to shared subspace inference with $\hat{s} = p$ because
this implies $VV^T = I_p$.  Although the risk is large for small
values of $\hat{s}$, as the shared subspace dimension increases to the
dimension of the feature space, that is $\hat{s} \rightarrow p$, the
risk for the shared subspace estimator quickly decreases.
Importantly, it is always true that rank($[U_1, ..., U_K]) \leq rK$ so
it can equivalently be assumed that the data were generated from a
shared subspace model with dimension $s = rK < p$.  As such, even when
there is little similarity between the eigenvectors from each group,
the shared subspace estimator with $\hat{s} = rK$ will perform well,
provided that we can identify a subspace, $\hat{V}\hat{V}^T$, that is
close to $\text{span}([U_1, ..., U_K])$. When
$\hat{V}\hat{V}^T = \text{span}([U_1, ..., U_K])$ exactly, shared
subspace estimation outperforms independent covariance estimation
(\ref{fig:sdimension}, dashed red line).

From this simulation, it is clear that correctly specifying the
dimension of the shared subspace is important for efficient covariance
estimation.  When the dimension of the shared subspace is too small,
we accrue higher risk.  The goodness of fit statistic,
$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$, can be used to identify when
a larger shared subspace is warranted.  When $\hat{s}$ is too small,
$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$ will be significantly smaller
than one for at least some of the groups, regardless of $\hat{V}$
(e.g. Figure \ref{fig:ratio-s5}).  When $\hat{s}$ is large enough, we
are able to use maximum marginal likelihood to identify a shared
subspace which preserves most of the variation in the data for all
groups (Figure \ref{fig:ratio-s20}).  Thus, for any estimated
subspace, the goodness of fit statistic can be used to identify the
groups that can be fairly compared on this subspace and whether we
would benefit from fitting a model with a larger value of $\hat{s}$.

\paragraph{Model Comparison and Rank Estimation:}

Clearly, correct specification for the rank of the shared subspace is
important for efficient inference.  So far in this section, we have
assumed that the group rank, $r$, and shared subspace dimension, $s$,
are fixed and known.  In practice this is not the case.  Prior to
fitting a model we should estimate these quantities.
\citet{Gavish2014} provide an asymptotically optimal (in mean squared
error) singular value threshold for low rank matrix recovery with
noisy data.  We use their rank estimator, which is a function of the
median singular value of the data matrix and the ratio
$\alpha_k =\frac{p}{n_k}$, to estimate $r$.  Although their estimator
was derived for individual covariance estimation, we found that,
Gavish and Donoho's estimator can also be applied to the pooled data
to estimate $s$.  Specifically, we concatenate the data from all
groups to create an $(\sum_k n_k) \times p$ dimensional matrix and
apply their rank estimator to this matrix to choose $s$.

Using these rank estimators, we conduct a simulation which
demonstrates the relative performance of shared subspace group
covariance estimation under different data generating models.  In
these simulations we assume the data have $p=200$ features, $r=2$
spikes, $\sigma^2_k=1$, and $n_k = 50$.  We fix the non-zero eigenvalues of
$\Psi_k$ to $(\lambda_1, \lambda_2) = (250, 25)$.  We consider three
different shared subspace data models: 1) a low dimensional shared
subspace model with $s=r=2$; 2) a model in which the spiked covariance
matrices from all groups are identical, e.g.
$\Sigma_k = \Sigma = U\Lambda U^T + \sigma^2I$; and 3) a full rank
shared subspace model with $s=p=200$.  We simulate 100 independent
datasets for each of these data generating mechanisms.

We fit these datasets using three different variants of the shared
subspace model.  For each of these fits we estimate $r$.  First, we
estimate a single spiked covariance matrix from the pooled data and
let $\hat{\Sigma}_k = \hat{\Sigma}$.  Second, we fit the full rank
shared subspace model.  This corresponds to a procedure in which we
estimate each spiked covariance matrix independently, since $s=p$
implies $VV^T = I_p$.  Finally, we use an ``adaptive'' shared subspace
estimator, in which we estimate both $s$, $r$ and $VV^T$.  In Table
\ref{table:groupLoss} we report the average Stein's risk and
corresponding 95\% loss intervals for the estimates derived from each
of these inferential models.

\begin{table}
\begin{center}
  \begin{tabular}{ l  l | c | c | c |}
    \multicolumn{2}{c}{} & \multicolumn{3}{c}{\textbf{Inferential Model}} \\
  \multicolumn{2}{c|}{}  & Adaptive & $\hat{\Sigma}_k=\hat{\Sigma}$
                                                           & $\hat{s} = p$ \\  \cline{2-5}
    \multirow{3}{*}{\rotatebox[origin=c]{90}{\textbf{Data Model}}} 
& $s=r=2$ & 0.8 (0.7, 0.9) & 2.1 (1.7, 2.6) & 3.0 (2.9, 3.2) \\ \cline{2-5}
   &   $s=r=2$, $\Sigma_k = \Sigma$ & 0.8 (0.7, 0.9) & 0.7 (0.6, 0.8) & 3.0 (2.9, 3.2)\\ \cline{2-5}
   &  $s=p=200$ & 7.1 (6.2, 8.0) & 138.2 (119, 153) & 3.0 (2.9, 3.2) \\ \cline{2-5}
  \end{tabular}
  \caption[Table caption text]{Stein's risk (and 95\% loss intervals)
    for different inferential models and data generating models with
    varying degrees of between-group covariance similarity.  For each
    of $K=10$ groups, we simulate data from three different types of
    shared subspace models.  For each of these models, $p=200$, $r=2$,
    $\sigma_k^2=1$ and $n_k=50$.  We also fit the data using three
    different shared subspace models: a model in which $s$, $r$ and
    $VV^T$ are all estimated from the data (``adaptive''), a spiked
    covariance model in which the covariance matrices from each group
    are assumed to be identical ($\hat{\Sigma}_k=\hat{\Sigma}$) and a
    model in which we assume the data do \emph{not} share a lower
    dimensional subspace across groups (i.e. $\hat{s} = p$). The estimators
    which most closely match the data generating model have the lowest
    risk (diagonal) but the adaptive estimator performs well relative
    to the alternative misspecified model.}
\label{table:groupLoss}
\end{center}
\end{table}

As expected, the estimates with the lowest risk are derived from the
inferential model that most closely match the data generating
specifications. However, the adaptive estimator has small risk under
model misspecification relative to the alternatives.  For example,
when $\Sigma_k = \Sigma$, the adaptive shared subspace estimator has
almost four times smaller risk than the full rank estimator, in which
each covariance matrix is estimated independently.  When the
data come from a model in which $s=p$, that is, the eigenvectors of $\Psi_k$
are generated uniformly from $\mathcal{V}_{p,r}$, the adaptive
estimator is over an order of magnitude better than the estimator
which assumes no differences between groups.  These results suggest
that empirical Bayes inference for $VV^T$ combined with the rank
estimation procedure suggested by \citet{Gavish2014} can be widely
applied to group covariance estimation because the estimator adapts to
the amount of similarity across groups.  Thus, shared subspace estimation
can be an especially appropriate choice when the similarity between
groups is not known a priori.  % In the simulations
% used to create Table \ref{table:groupLoss}, we typically estimated $s$
% to be smaller than $rK$ for the data generated under ``full rank''
% shared subspace model data.  This explains the slightly higher risk
% for the shared subspace model. With improved estimators for $s$, the
% shared subspace model should be even more competitive.

% Ultimately, in many real world examples, we are more interested in
% finding a useful space to interpret and compare differences between
% covariance matrices than we are in finding the lowest-loss estimator
% for the full covariance matrices.  In this context, the relevant
% quantity is often an estimate of the projected covariance, $V^T\Sigma
% V$.

% In Figure \ref{fig:lossBoxplot} we depict boxplots for
% the distribution of losses,
% $L_S(\hat{V}^T\Sigma_k\hat{V}, \hat{V}^T\hat{\Sigma}_k\hat{V})$, on
% simulated data.  In the left boxplot the data come from a
% 2-dimensional shared subspace model and on the right plot the data
% come from a 10-dimensional shared subspace model.  For both boxplots,
% estimates $\hat{V}^T\hat{\Sigma}_k\hat{V}$ are derived from assumed
% 2-dimensional model. This figure demonstrates that the loss for
% estimating the projected covariance matrix is not significantly larger
% when we underestimate the dimension of the true shared subspace.

\section{Analysis of Biological Data}
\label{sec:app}

In this Section we apply shared subspace spiked covariance models to data in two
different biological examples.  In the first, we compare gene
expression data from juveniles with different subtypes of leukemia.  Even
after removing mean effects, our results indicate significant differences
in covariance matrices between groups.  In the second example, we
compare covariance matrices in a metabolomic analysis of \textit{Drosophila} data.  In
this example, after controlling for mean effects due to fly age, gender and
genotype, we find little compelling evidence of differences in metabolite
covariances between groups.

In these analyses we employ a shared subspace model in
which we define the $s$ by $s$ matrix $\Psi_k$ as
%
\begin{equation}
\Psi_k =\left( \begin{array}{cc}
O_k\Lambda_kO_k^T & 0  \\
0 & D  \end{array} \right)
\end{equation}
%
For the following analyses $O_k\Lambda_kO_k^T$ is a rank $r=2$ matrix
and $D$ is an $(s-r)$-dimensional diagonal matrix.  We write
$V = [V_1, V_2]$, with $V_1 \in \mathcal{V}_{p,r}$ as the basis for a
$r$-dimensional shared subspace that explains the differences between
groups and $V_2 \in \mathcal{V}_{p,(s-r)}$ corresponding to the remaining $(s-r)$
eigenvectors common to all groups.  Importantly, we make no assumptions about the
magnitude of the eigenvalues $\Lambda_k$ relative to the eigenvalues
$D$, so that the largest eigenvectors of $\Psi_k$ may correspond to
the eigenvalues from columns of $V_2$.

We find this formulation useful because in real world analyses,
differences between groups may not manifest themselves in the first
principle components.  For instance, in genetic analyses, there may be
large between subject variability common to all groups which is
unrelated to how those subjects are grouped. Differences between
groups may manifest themselves in the smaller principle
components. This formulation allows a complete pooling of several
eigenvectors, yet a space to identify and compare relevant
differences.  Using this framework we can easily visualize and compare the
largest differences between groups in a two dimensional space while
completely pooling remaining anisotropic variability that is not
specific to the grouping.

\paragraph{Analysis of Gene Expression Data:}

We demonstrate the utility of the shared subspace covariance
estimator for exploring differences in the covariability of gene
expression levels in young adults with different subtypes of pediatric
acute lymphoblastic leukemia \citep{Yeoh2002}.  The raw data consists
of gene expression levels for thousands of genes in seven different
subtypes of leukemia: BCR-ABL, E2A-PBX1, Hyperdip50, MLL, T-ALL,
TEL-AML1 and a seventh group for unidentified subtypes (``Others'').  The groups
have corresponding sample sizes of $n = (15, 27, 64, 20, 43, 79, 79)$.
Although there are over 12,000 genes in the dataset, the vast majority of expression
levels are missing.  Thus, we restrict our attention to the genes for
which less than half of the values are missing and use Amelia, a
software package for missing value imputation, to fill in the
remaining missing values \citep{Amelia}.  After restricting the data
in this way, $p=3124$ genes remain.  Prior to analysis, we demean both
the rows and columns of the gene expression levels in each group.

In Figure \ref{fig:leukemia} we plot the results of our analysis.
Panel \ref{fig:leukemiaRatio} shows that over $60\%$ of the variance
can be explained by the  shared subspace for all
groups, with over 80\% of the variance explained in  4 of the 6 groups.  Panel
\ref{fig:leukemiaPosterior} reflects some significant differences in
the posterior distribution of eigenvalues and eigenvectors between
groups.  The $x$-axis corresponds to the orientation of the principle
axis of $\Psi_k$ on $\hat{V}_1$ and the $y$-axis corresponds to the log ratio of
the first two eigenvalues.  Several groups have significantly
different orientations for the first principle component of the
projected covariance matrix, as well as differences in the ratio of
eigenvalues.  For instance, the subtype E2A-PBX1 has a larger
eigenvalue ratio which reflects a more correlated distribution on this
subspace.  The posterior regions for the TEL-AML1 and Hyperdip50
groups are almost entirely overlapping, which suggests there is little
detectable difference between their covariance structures.  Note that
when the eigenvalue ratio is close to one (as is the case for the
``other'' subtype), the distribution of expression values on the
subspace is nearly spherical and thus the orientation of the primary
eigenvector is at best weakly identifiable.  This is reflected in the
wide posterior range of orientations of the first principle component.

Further intuition about the differences in covariances between groups
can be understood in the biplot in Figure \ref{fig:leukemiaBiplot}.
Here, we plot the contours of the covariance matrices for three
leukemia subtypes. The 1\% of genes with the largest loadings on the
first two columns of $\hat{V}_1$ are indicated with black shapes and the
remaining loadings with light grey dots.  The genes with the largest
loadings are clustered by quadrant and listed in the corresponding
table.  Even though we remove all mean-level differences in gene
expressions between groups, we are still able to identify genes with
known connections to cancer and leukemia.  As one example, MYC (top
right) is an oncogene with well established association to many
cancers \citep{Dang2012}.

\begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{Figs/leukemiaRatio-07-01}
        \caption{$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$}
        \label{fig:leukemiaRatio}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{Figs/leukemiaPosterior-07-01}
        \caption{95\% Posterior Regions }
        \label{fig:leukemiaPosterior}
      \end{subfigure}
      \caption{a) Goodness of shared subspace fit for each of the
        seven Leukemia groups.  Over 60\% of the variance is explained
        on the shared subspace in each of the seven groups, with
        more than 80\% explained in most groups.  b) 95\% posterior
        regions for the eigenvalue ratio and primary eigenvector
        orientation.  Regions for some pairs of groups are disjoint,
        suggesting significant differences in the projected covariance
        matrices.  For other groups (e.g. Hyperdip50 and TEL-AML1)
        overlap in the posterior regions indicate that differences are
        not detectable on this subspace. }
\label{fig:leukemia}
\end{figure}

In this figure, genes that fall in the upper right quadrant have
positive loadings in both columns of $\hat{V}_1$. Genes in the upper left
corner of the figure have positive loadings on the first column of $\hat{V}_1$
but negative loadings on the second column of $\hat{V}_1$.  The principle axis
for a group aligns in the direction of genes that exhibit the largest
variability in that group.  Genes which lie in a direction orthogonal
to the principle axis exhibit reduced variability, relative to the
other groups. As an example, genes in the upper left corner of this
plot (triangles) exhibit large, positively correlated variability in
the Hyperdip50 group.  In this same group, there is reduced
variability, relative to the other groups, among the genes in the
upper right corner of the plot (squares), since this cluster of genes lie in
a direction nearly orthogonal to the principle axis.  In contrast, the
BCR-ABL group is aligned primarily with the second column of $\hat{V}_1$,
which means that the genes indicated by squares and circles vary
significantly in this group.  Genes in the square group are
anti-correlated with those in the circle group, since their
loadings have opposite signs on the second column of $\hat{V}_1$.

  \begin{figure}[!ht]
    \centering
    \includegraphics[width=0.5\textwidth]{Figs/leukemia-biplot-05-13}
    \qquad
\raisebox{1.25\height}{
\footnotesize
\input{Figs/biplot.tab}
}
\caption{Left) Variant of a biplot with contours for three leukemia
  subtypes and the loadings for each gene on the first two columns of
  $\hat{V}$.  All gene's loadings are displayed in light gray, and the
  top 1\% of genes with with the largest magnitude loadings are
  displayed as either a triangle, square or circle depending on which
  quadrant they lie in.  Right) List of the gene's with the largest
  loadings, grouped by quadrant. }
\label{fig:leukemiaBiplot}
  \end{figure}

\paragraph{Analysis of Metabolomic Data:}

Next, we briefly discuss an example in which we could not readily identify
significant covariance differences between groups.  Here, we apply
shared subspace group covariance estimation to a metabolomic data
analysis on fly aging \citep{Hoffman2014}.  We bin the data to include
groups of flies less than 10 days (young), between 10 and 51 days
(middle) and greater than 51 days (old), and further split by
sex. The sample sizes range from 43 to 59 flies per group.  We
analyze metabolomic data corresponding to metabolites with 3714
mass-charge ratios.  After removing mean effects due to age and
sex, we fit the shared subspace estimator and compare covariances
between groups.  We identify a subspace that explains almost all of
the variability in the first few components of $\Sigma_k$ for all
groups (Figure \ref{fig:dmelanRatio}) but see little evidence for
differences in metabolite covariances on this subspace, as evidenced
by the large overlap in posterior distributions for each group (Figure
\ref{fig:dmelanPosterior}). This could be indicative of a large amount
of variation that is common across all groups for the first $s$
principle components. Differences in covariation across groups likely
manifest themselves in the smaller principle components but are too
small to detect given the sample sizes.

\begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{Figs/dmelanRatio-04-28}
        \caption{$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$}
        \label{fig:dmelanRatio}
      \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.45\textwidth}r
        \includegraphics[width=\textwidth]{Figs/dmelanPosterior-07-01}
        \caption{95\% Posterior Regions}
        \label{fig:dmelanPosterior}
    \end{subfigure}
    \caption{a) Goodness of fit is close to one for all groups
      suggesting that the inferred subspace preserves most of the
      variability across groups. b) There is significant overlap in
      the posterior distribution of the log eigenvalue ratio and primary
      eigenvector orientation across groups. Thus, there is little
      evidence of significant differences in the covariance matrices
      between groups on this inferred subspace.  }
\end{figure}

\section{Discussion}

In this paper, we proposed a class of models for estimating and
comparing differences in covariance matrices across multiple groups on
a common low dimensional subspace.  We described an empirical Bayes
algorithm for estimating this common subspace and a Gibbs sampler for
inferring the projected covariance matrices and their associated
uncertainty.  Estimates of both the shared subspace and the projected
covariance matrices can both be useful summaries of the data.  For
example, with the biological data, the shared subspace highlights the
full set of genes or metabolites that are correlated across group.
Differences between group covariance matrices can be understood in
terms of differences in these sets of correlated molecules.  In the applied analysis, we demonstrated how
we can use these notions to visualize and contrast the posterior
distributions of covariance matrices projected onto a particular
subspace.

In simulation, we showed that the shared subspace model can still be a
reasonable choice for modeling multi-group covariance matrices even
when the groups may be largely dissimilar.  When there is little
similarity between groups, the shared subspace model can still be appropriate as
long as the dimension of the shared subspace is large enough.
However, selecting the rank of the shared subspace remains a practical
challenge.  Although we propose a useful heuristic for choosing the
dimension of the shared subspace based on the rank selection
estimators of \citet{Gavish2014}, a more principled approach is
warranted.  Improved rank estimators would further improve the
performance of the adaptive shared subspace estimator discussed in
Section \ref{sec:simulation}.

It is also a challenging problem to estimate the ``best'' subspace
once the rank of the space is specified.  We used maximum marginal
likelihood to estimate $VV^T$ and then use MCMC to
infer $\Psi_k$.  By focusing on group differences for $\Psi_k$ on a
\emph{fixed} subspace, it is much simpler to interpret similarities
and differences.  Nevertheless, full uncertainty quantification for
$VV^T$ can be desirable.  We found MCMC inference for $VV^T$ to be
challenging for the problems considered in this paper and leave it for
future work to develop an efficient fully Bayesian approach for
estimating the joint posterior of $VV^T$ and $\Psi_k$.  Recently
developed Markov chain Monte Carlo algorithms, like Riemmanian
manifold Hamilton Monte Carlo, which can exploit the geometry of the
Grassmann manifold, may be useful here \citep{Byrne2013,
  Girolami2011}.  It may also be possible, though computationally
intensive, to jointly estimate the $s$ and $VV^T$ using for instance,
a reversible-jump MCMC algorithm.

% There are many different prior specifications for $\Psi_k$ and
% $\sigma_k^2$ that we could consider for estimating posterior
% uncertainty.  One simple approach would be to use the same prior
% specification used in Section \ref{sec:em}.  That is, we can use
% Jeffreys' prior for $\frac{1}{\sigma_k^2}$ and
% $M_k^{-1} = (\Psi_k + I) ^{-1}/\sigma_k^2$.  In fact, it is easy to
% show that the Gamma distribution for $\frac{1}{\sigma_k^2}$ and the
% Wishart distribution for $M_k^{-1}$ are conjugate priors for the
% respective quantities.  From posterior samples of $M_k$ and
% $\sigma_k^2$ we can easily recover samples of $\Psi_k$.
% Unfortunately, if $M_k^{-1}$ has a Wishart distribution, the implied
% support of $\Psi_k$ is not constrained to the cone of positive
% semi-definite matrices.  Direct prior specifications for $\Psi_k$ are
% more appealing, but many obvious choices, like the inverse-Wishart
% prior distribution for $\Psi_k$ do not lead to simple posterior
% distributions.

Fundamentally, our approach is quite general and can be integrated
with existing approaches for multi-group covariance
estimation.  In particular, we can incorporate additional shrinkage on
the projected covariance matrices $\Psi_k$.  As in \citet{Hoff2009} we
can employ non-uniform Bingham prior distributions for the
eigenvectors of $\Psi_k$ or we can model $\Psi_k$ as a function of continuous
covariates as in \citet{Yin2010} and \citet{Hoff2011}.  Alternatively, we can
summarize the estimated covariance matrices by thresholding entries of
the precision matrix, $\Psi_k^{-1}$ to visualize differences between
groups using a graphical model \citep{Meinshausen2006}.  The specifics
of the problem at hand should dictate which shrinkage models are
appropriate, but the shared subspace assumption can be useful in a
wide range of analyses, especially when the number of features is very
large.

\bibliographystyle{plainnat}
\bibliography{refs.bib}

%\section{Appendix}

\end{document}